---
title: "12_transpose_counts"
output: html_document
date: "2024-08-27"
---
# Transpose counts
This script reads in counts from stage 11 (which are written with each person as a different column, and each gene as a different row) and transposes it so the rows are people, and the columns are each gene. This still works on a per chromosome level as the tables would be too big if they spanned all chromosomes. This stage was run using a mem2_ssd2_v2_x16 instance and took 3 hours, costing £2.

## Load packages
```{r}
library(data.table)
```

# Transpose table of burden and save out
```{r}
# Make a list of chromosomes, variant classes, and ACs to process (ensure names match up to table names)
chromosomes <- 1:23
variant_classes <- c("REVEL75_Miss", "PTV", "synonymous")
allele_counts <- '5orless'   

# Loop over each chromosome
for (chr in chromosomes) {
    for (var in variant_classes) {
        for (AC in allele_counts) {
          if (chr == 1) {
            halves <- c("first_half", "second_half")
            for (half in halves) {
                    # Construct the file name for the current half
                    input_filename <- paste0('./chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    
                    # Download the file to your project using dx download
                    download_command <- paste('dx download ', shQuote(input_filename))
                    system(download_command)

                    # Read the table for the current half using fread
                    input_filename <- paste0('chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
                     # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
                    counts=subset(counts, !is.na(counts$gene_id)) 
                    
                    # Transpose the table using melt and dcast
                    transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
                    transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")

                    # Write the transposed table out
                    output_filename <- paste0('transposed_chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    fwrite(transposed_counts, output_filename, sep = "\t")

                    # Upload the file to your project using dx upload
                    upload_command <- paste('dx upload', shQuote(output_filename))
                    system(upload_command)
                }
            }
          else {
            # Construct the file name 
            input_filename <- paste0('./chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')

            # Download the file to your project using dx download
            download_command <- paste('dx download ', shQuote(input_filename))
            system(download_command)
            
            # Read the table for the current chromosome using fread and transpos
            ## Using fread from data table really speeds up the process of reading in 
            input_filename <- paste0('chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
            # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
            counts=subset(counts, !is.na(counts$gene_id)) 
            
            # Transpose the table using melt and dcast:
            ## Melt takes table from wide format into a v long format, where each row represents a single observation (so a single gene and a single sample)
            ### This transposed table is really big so you need a fairly high mem session for this.
            transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
            ## dcast then takes the dataset in this long format and casts it back to a wide format, but this time with sample id as rows and genes as columns
            transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")
            
            # Write the transposed table out 
            output_filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            fwrite(transposed_counts, output_filename, sep = "\t")

            # Upload the file to your project using dx upload
            upload_command <- paste('dx upload', shQuote(output_filename),'--destination ./Variant_counts/transposed_counts/')
            system(upload_command)
          } 
        }
    }
}
          
```
