---
title: "16_gene_set_regressions"
output: html_document
date: "2024-10-01"
---

# Gene set regressions
This script reads in phenotypic data and tables of counts per gene set, and then runs regressions on these data. This script runs for <1 hour minutes on a mem1_ssd1_v2_x8 RStudio instance, costing £0.20.
# Load libraries
```{r}
library(data.table)
library(dplyr)
library(purrr)
install.packages('broom')
library(broom)
library(ggplot2)
```

# Prep data
## Read in cognitive phenotypes
```{r}
system('dx download ./phenotype_cohorts/processed_cognition/processed_cog_tests_inc_g_Feb24.csv')
phen=read.csv('processed_cog_tests_inc_g_Feb24.csv', header=T)

```

## Ancestries
```{r}
system('dx download ./genetic_ancestries/app15175_ids/full_dataset_inferred_ancestries.csv')
ancs_wrongid=read.csv('full_dataset_inferred_ancestries.csv', header=T)
system('dx download ./genetic_ancestries/app15175_ids/PCAs_*')
EUR_PCs_wrongid=read.table('PCAs_EUR_only.tsv', header=T)
AAC_PCs_wrongid=read.table('PCAs_AAC_only.tsv', header=T)
EAS_PCs_wrongid=read.table('PCAs_EAS_only.tsv', header=T)
LAT_PCs_wrongid=read.table('PCAs_LAT_only.tsv', header=T)
SAS_PCs_wrongid=read.table('PCAs_SAS_only.tsv', header=T)
SSA_PCs_wrongid=read.table('PCAs_SSA_only.tsv', header=T)
ADM_PCs_wrongid=read.table('PCAs_ADM_only.tsv', header=T)

system('dx download ./phenotype_cohorts/link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt')
link_ids=read.table('link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt', header=T)
ancs=merge(ancs_wrongid, link_ids, by.x='FID', by.y='app15175')
ancs=merge(ancs, EUR_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, AAC_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, EAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, LAT_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SSA_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, ADM_PCs_wrongid, by.x ='FID', by.y='s', all=T)

# Write out now have correct IDs
write.csv(ancs, 'ancestries_PCs.csv')
system('dx upload ancestries_PCs.csv ./genetic_ancestries/')

# Merge 
phen=merge(phen, ancs, by.x = 'eid', by.y='eid_13310')

```

## QC
Filter to those passing the final QC stage (sex imputation)
```{r}
system('dx download ./WES_QC/discordant_sex_ids.csv')
ids_to_rm=read.csv('discordant_sex_ids.csv')
ids_to_rm$remove=T


phen=merge(phen, ids_to_rm, by.x = 'eid', by.y='x', all=T)
phen=subset(phen, is.na(phen$remove)) 

```

## Covariates
Read in covariates (and prep for use). These phenotypes were selected in cohort browser and written out using the Table Exporter tool. 
```{r}
system('dx download ./phenotype_cohorts/covars_participant.csv')
covars=read.csv('covars_participant.csv', header=T)

covars$seq_batch<-covars$p32050
covars$AC_attended<-covars$p54_i0
covars$sex_1male[covars$p31 == 'Male']<-1
covars$sex_1male[covars$p31 == 'Female']<-2
covars$standardised_age <- (covars$p21022 - mean(covars$p21022, na.rm = TRUE)) / sd(covars$p21022, na.rm = TRUE)
covars$standardised_age_squared<-covars$standardised_age*covars$standardised_age
covars$sex_standardised_age<-covars$sex_1male * covars$standardised_age
covars$sex_standardised_age_squared<-covars$sex_1male * covars$standardised_age_squared


phen=merge(phen, covars, by.x='eid', by.y='eid')


```

## Diagnoses 
These phenotypes were selected in cohort browser and written out using the Table Exporter tool. 
```{r}
system('dx download ./phenotypes_cog_diagnoses_participation_death.csv')
diagnoses=read.csv('phenotypes_cog_diagnoses_participation_death.csv', header=T)

# Make subset of people diagnosed with SZ - to run these regressions in if powered? 
sz_diagnosed=subset(diagnoses, diagnoses$SZ==T) 
sz_diagnosed_eur_only=subset(sz_diagnosed, sz_diagnosed$EUR>0.8)

# Remove people with SZ, ID and/or ASD 
diagnoses=subset(diagnoses, diagnoses$SZ==F & diagnoses$ASD==F & diagnoses$intellectual_disability==F) 
diagnoses$no_diagnoses<-T
diagnoses=diagnoses[,c(3,174)]

phen=merge(phen, diagnoses, by='eid') 
```

## Phenotype counts across ancs
```{r}
all_ancs=phen
EUR_anc_ids=subset(phen, phen$EUR>0.8)
AAC_anc_ids=subset(phen, phen$AAC>0.8)
EAS_anc_ids=subset(phen, phen$EAS>0.8)
LAT_anc_ids=subset(phen, phen$LAT>0.8)
SAS_anc_ids=subset(phen, phen$SAS>0.8)
SSA_anc_ids=subset(phen, phen$SSA>0.8)
Admixed_anc_ids=subset(phen, (phen$AAC<=0.8 & phen$EAS<=0.8 & 
                              phen$EUR<=0.8 & phen$LAT<=0.8 & 
                              phen$SAS<=0.8 & phen$SSA<=0.8))



# With g
a=subset(all_ancs, !is.na(all_ancs$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm))

# Numeric memory 
a=subset(all_ancs, !is.na(all_ancs$Numeric_memory_online_maximum_digits_remembered_Zscore))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$Numeric_memory_online_maximum_digits_remembered_Zscore))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$Numeric_memory_online_maximum_digits_remembered_Zscore))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$Numeric_memory_online_maximum_digits_remembered_Zscore))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$Numeric_memory_online_maximum_digits_remembered_Zscore))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$Numeric_memory_online_maximum_digits_remembered_Zscore))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$Numeric_memory_online_maximum_digits_remembered_Zscore))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$Numeric_memory_online_maximum_digits_remembered_Zscore))

# Pairs matching
a=subset(all_ancs, !is.na(all_ancs$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))

# Reaction time
a=subset(all_ancs, !is.na(all_ancs$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))


# TMTB
a=subset(all_ancs, !is.na(all_ancs$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))

# FI
a=subset(all_ancs, !is.na(all_ancs$FI_AC1_n_correct_answers_Zscore))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$FI_AC1_n_correct_answers_Zscore))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$FI_AC1_n_correct_answers_Zscore))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$FI_AC1_n_correct_answers_Zscore))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$FI_AC1_n_correct_answers_Zscore))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$FI_AC1_n_correct_answers_Zscore))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$FI_AC1_n_correct_answers_Zscore))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$FI_AC1_n_correct_answers_Zscore))

# SDS
a=subset(all_ancs, !is.na(all_ancs$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))


# TMTA
a=subset(all_ancs, !is.na(all_ancs$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))

# Prospective mem
a=subset(all_ancs, !is.na(all_ancs$prospective_mem_AC1))
a=subset(EUR_anc_ids, !is.na(EUR_anc_ids$prospective_mem_AC1))
a=subset(SAS_anc_ids, !is.na(SAS_anc_ids$prospective_mem_AC1))
a=subset(AAC_anc_ids, !is.na(AAC_anc_ids$prospective_mem_AC1))
a=subset(SSA_anc_ids, !is.na(SSA_anc_ids$prospective_mem_AC1))
a=subset(EAS_anc_ids, !is.na(EAS_anc_ids$prospective_mem_AC1))
a=subset(Admixed_anc_ids, !is.na(Admixed_anc_ids$prospective_mem_AC1))
a=subset(LAT_anc_ids, !is.na(LAT_anc_ids$prospective_mem_AC1))

# Other gs 
a=subset(all_ancs, !is.na(all_ancs$g_LOO_RT_TMTB_NM_zscore_outliersrm))
a=subset(all_ancs, !is.na(all_ancs$g_LOO_RT_NM_PM_zscore_outliersrm))
a=subset(all_ancs, !is.na(all_ancs$g_LOO_TMTB_NM_PM_zscore_outliersrm))
a=subset(all_ancs, !is.na(all_ancs$g_LOO_TMTB_NM_RT_zscore_outliersrm))
a=subset(all_ancs, !is.na(all_ancs$g_inc_FI_zscore_outliersrm))
a=subset(all_ancs, !is.na(all_ancs$g_UKB5_FRD_zscore_outliersrm.y))

```

# Correlations
## Between cog tests
```{r}
cog_tests_for_correlations<- all_ancs[,c('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm', 
                                         'Numeric_memory_online_maximum_digits_remembered_Zscore',  
                                         'Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm',
                                         'Reaction_time_AC1_mean_time_taken_Zscore_outliersrm',
                                         'TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm',
                                         'FI_AC1_n_correct_answers_Zscore',
                                         'SDS_online1_n_correct_cons_attempts_Zscore_outliersrm',
                                         'TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm'
                                         )]

correlations <- cor(cog_tests_for_correlations, use="pairwise.complete.obs")
```

## Of different gs 
```{r}
gs_for_correlations<- all_ancs[,c('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm', 
                                         'g_LOO_RT_TMTB_NM_zscore_outliersrm',  
                                         'g_LOO_RT_NM_PM_zscore_outliersrm',
                                         'g_LOO_TMTB_NM_PM_zscore_outliersrm',
                                         'g_LOO_TMTB_NM_RT_zscore_outliersrm',
                                         'g_inc_FI_zscore_outliersrm',
                                         'g_UKB5_FRD_zscore_outliersrm.y'
                                         )]

correlations <- cor(gs_for_correlations, use="pairwise.complete.obs")
```


# Get counts
```{r}
AC='5orless'
system('dx download ./Variant_counts/Gene_sets/*_AC5orless_*.csv --overwrite')

# REMOVE SINGLE GENE TABLES HERE!!
file_list <- list.files(path = "./", pattern = "*_AC5orless_counts.csv", full.names = TRUE)

# Read, drop first column (V1), and merge each file
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames


all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 
AAC_ancs=subset(all_ancs, all_ancs$AAC>0.8)
EAS_ancs=subset(all_ancs, all_ancs$EAS>0.8)
LAT_ancs=subset(all_ancs, all_ancs$LAT>0.8)
SAS_ancs=subset(all_ancs, all_ancs$SAS>0.8)
SSA_ancs=subset(all_ancs, all_ancs$SSA>0.8)
ADM_ancs=subset(all_ancs, (all_ancs$AAC<=0.8 & all_ancs$EAS<=0.8 & 
                              all_ancs$EUR<=0.8 & all_ancs$LAT<=0.8 & 
                              all_ancs$SAS<=0.8 & all_ancs$SSA<=0.8))



```


## EUR ancs regression (MAC5)
### Set covars 
```{r}
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' # SSA removed here as no one defined as EUR anc has a non-zero SSA probability. 

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
```

### Prep results table 
```{r}
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)
```

### Distribution of RCVs
#### In LoFi genes 
Supplementary Table.
```{r}
a=subset(EU_ancs, !is.na(EU_ancs$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm)) #75,188

a %>% count(PTV_AC5orless_pLI_0.9_total)
a %>% count(REVEL75_Miss_AC5orless_pLI_0.9_total)
a %>% count(synonymous_AC5orless_pLI_0.9_total)

b = subset(a, a$synonymous_AC5orless_pLI_0.9_total > 5)

```


#### compared to g 
Supplementary Figure 3
```{r}
# Combine data for all variant classes into a single data frame
summaryPTVS <- a %>%
  group_by(PTV_AC5orless_pLI_0.9_total) %>%
  summarise(
    mean_g = mean(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm),
    se = sd(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm) / sqrt(n()),
    variant_class = "PTVs"
  )

summaryMissPathogenic <- a %>%
  group_by(REVEL75_Miss_AC5orless_pLI_0.9_total) %>%
  summarise(
    mean_g = mean(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm),
    se = sd(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm) / sqrt(n()),
    variant_class = "Deleterious missense"
  )

summarySynonymous <- a %>%
  group_by(synonymous_AC5orless_pLI_0.9_total) %>%
  summarise(
    mean_g = mean(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm),
    se = sd(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm) / sqrt(n()),
    variant_class = "Synonymous"
  )

# Combine all summaries into one data frame
combined_summary <- bind_rows(
  summaryPTVS %>% rename(n_variants = PTV_AC5orless_pLI_0.9_total),
  summaryMissPathogenic %>% rename(n_variants = REVEL75_Miss_AC5orless_pLI_0.9_total),
  summarySynonymous %>% rename(n_variants = synonymous_AC5orless_pLI_0.9_total)
)

# Reorder factor levels for variant_class
combined_summary$variant_class <- factor(
  combined_summary$variant_class,
  levels = c("PTVs", "Deleterious missense", "Synonymous")
)

# Drop synonymous here as we dont want to plot this 
combined_summary=subset(combined_summary, combined_summary$variant_class!='Synonymous')

# Plot all variant classes in one graph
tiff("combined_variant_classes_by_g_means_SEs.tiff", units="in", width=15, height=10, res=400)
ggplot(combined_summary, aes(x = n_variants, y = mean_g, color = variant_class)) +
  geom_point(stat = "identity", position = position_dodge(width = 0.6), size = 6) +
  geom_errorbar(aes(ymin = mean_g - se, ymax = mean_g + se),position = position_dodge(width = 0.6), width = 0.2) +
  labs(x = '\n n rare variants in LoFi genes carried \n', 
       y = expression("Mean " * italic("g") * " score"),
       color = "Variant Class") +
  theme_minimal() +
  geom_hline(yintercept = 0, linetype = 'dashed', color = 'gray') +
  theme(text = element_text(size = 35))+
  xlim(-0.5,3.5)+
  ylim(-0.35, 0.05)+
  theme(legend.position="bottom")
dev.off()
````


### Rates of RCVs across ancestries 
Supplementary Table 4
```{r}
all_ancs$anc<-'UNK'
all_ancs$anc[all_ancs$AAC>0.8]<- 'AAC'
all_ancs$anc[all_ancs$EAS>0.8]<- 'EAS'
all_ancs$anc[all_ancs$EUR>0.8]<- 'EUR'
all_ancs$anc[all_ancs$LAT>0.8]<- 'LAT'
all_ancs$anc[all_ancs$SAS>0.8]<- 'SAS'
all_ancs$anc[all_ancs$SSA>0.8]<- 'SSA'

all_ancs%>%count(anc)

summary_ancs <- all_ancs %>%
  group_by(anc) %>%
  summarise(
    #Exome wide
    mean_EW_PTVs = mean(PTV_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_PTVs = sd(PTV_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    mean_EW_pathogenic_missense = mean(REVEL75_Miss_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_pathogenic_missense = sd(REVEL75_Miss_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    mean_EW_intermediate_missense = mean(REVEL75to50_Miss_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_intermediate_missense = sd(REVEL75to50_Miss_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    mean_EW_benign_missense = mean(REVEL50orless_Miss_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_benign_missense = sd(REVEL50orless_Miss_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    mean_EW_synonymous = mean(synonymous_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_synonymous = sd(synonymous_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    # LoFi
    mean_LoFi_PTVs = mean(PTV_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_PTVs = sd(PTV_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n()),
    
    mean_LoFi_pathogenic_missense = mean(REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_pathogenic_missense = sd(REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n()),
    
    mean_LoFi_intermediate_missense = mean(REVEL75to50_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_intermediate_missense = sd(REVEL75to50_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n()),
    
    mean_LoFi_benign_missense = mean(REVEL50orless_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_benign_missense = sd(REVEL50orless_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n()),
    
    mean_LoFi_synonymous = mean(synonymous_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_synonymous = sd(synonymous_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n())
  )

  write.csv(summary_ancs, 'summary_ratesacrossancs.csv')

summary_all_ancs <- all_ancs %>%
  summarise(
    #Exome wide
    mean_EW_PTVs = mean(PTV_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_PTVs = sd(PTV_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    mean_EW_pathogenic_missense = mean(REVEL75_Miss_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_pathogenic_missense = sd(REVEL75_Miss_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    mean_EW_intermediate_missense = mean(REVEL75to50_Miss_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_intermediate_missense = sd(REVEL75to50_Miss_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    mean_EW_benign_missense = mean(REVEL50orless_Miss_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_benign_missense = sd(REVEL50orless_Miss_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    mean_EW_synonymous = mean(synonymous_AC5orless_exome_wide_total, na.rm = TRUE),
    se_EW_synonymous = sd(synonymous_AC5orless_exome_wide_total, na.rm = TRUE) / sqrt(n()),
    
    # LoFi
    mean_LoFi_PTVs = mean(PTV_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_PTVs = sd(PTV_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n()),
    
    mean_LoFi_pathogenic_missense = mean(REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_pathogenic_missense = sd(REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n()),
    
    mean_LoFi_intermediate_missense = mean(REVEL75to50_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_intermediate_missense = sd(REVEL75to50_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n()),
    
    mean_LoFi_benign_missense = mean(REVEL50orless_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_benign_missense = sd(REVEL50orless_Miss_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n()),
    
    mean_LoFi_synonymous = mean(synonymous_AC5orless_pLI_0.9_total, na.rm = TRUE),
    se_LoFi_synonymous = sd(synonymous_AC5orless_pLI_0.9_total, na.rm = TRUE) / sqrt(n())
  )


```



### Regress!!
#### Non-synonymous vars 
##### Set variables 
```{r}
sets <- list('exome_wide',
             'pLI_0.9','non_constrained_pLI_lessthan0.9',
             'brain_expressed_LoFi', 'brain_expressed_LoFtolerant',
             'brain_expressed','non_brain_expressed',
             'non_brain_expressed_LoFi','non_brain_expressed_LoFtolerant',
             'SZ_GWAS_all', 'SZ_GWAS_closest', 'SZ_GWAS_credible_causal_US','non_prioritised_GWAS', 
             'Brain_expressed_minus_credible_causal_GWAS','Brain_expressed_minus_SMR_prioritised_GWAS',
             'Finemap_prioritised_GWAS','SMR_prioritised_GWAS',
             'LoFi_minus_credible_causal_GWAS', 
             'NDD_CNVs_US', 'SZ_CNVs_US',
             'LoFi_in_NDD_CNVs_US', 'LoFi_in_SZ_CNVs_US',
             'LoFtolerant_in_NDD_CNVs_US','LoFtolerant_in_SZ_CNVs_US',
             'LoFi_minus_NDD_CNVs_US','LoFi_minus_SZ_CNVs_US',
             'SCHEMA_FDR_US', 'LoFi_and_SCHEMA_FDR_US',
             'LoFtol_and_SCHEMA_FDR_US', 'LoFi_minus_SCHEMA_FDR_US',
             'BE_and_SCHEMA_FDR_US', 'BE_minus_SCHEMA_FDR_US')

var_types<-list('PTV', 'REVEL75_Miss')
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

```

### Regress main sets!
```{r}
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

```


### Synonymous vars  
#### Set variables 
```{r}
var_types<-list('synonymous')

```

#### Regress
```{r}
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```



###  Sex stratified regression
#### Males 
##### Non-synonymous vars 
###### Set variables 
```{r}
sets <- list('pLI_0.9')
var_types<-list('PTV', 'REVEL75_Miss')

alldata=EU_ancs
EU_ancs=subset(EU_ancs, EU_ancs$sex_1male==1)
```

###### Regress
```{r}
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_males'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

```


##### Synonymous vars  
###### Set variables 
```{r}
var_types<-list('synonymous')

```

###### Regress
```{r}
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_males'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

#### Females 
##### Non-synonymous vars 
###### Set variables 
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

EU_ancs=alldata
EU_ancs=subset(EU_ancs, EU_ancs$sex_1male==2)
```

###### Regress
```{r}
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_females'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

```


#### Synonymous vars  
##### Set variables 
```{r}
var_types<-list('synonymous')

```

##### Regress
```{r}
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_females'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```



#### Write out results in case of crashes 
At this stage all ancs only but when EU PCs are ready, rerun and write out all results. 
```{r}
write.csv(results, 'results_g_regressions_AC5orless_diagnosesrm_incSE_Eurancs_Feb25_midway.csv')
system('dx upload results_g_regressions_AC5orless_diagnosesrm_incSE_Eurancs_Feb25_midway.csv --destination ./Variant_counts/Gene_sets/')
```


### Regress covarying for cognitive tests!! 
Just do this for lofi genes!
```{r}
sets <- list('pLI_0.9')

EU_ancs=alldata

```

#### g as outcome 
```{r}
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

```
##### Covarying for FI 
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS, "+ FI_AC1_n_correct_answers_Zscore"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$FI_AC1_n_correct_answers_Zscore))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_FI_AC1_n_correct_answers_Zscore'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn, "+ FI_AC1_n_correct_answers_Zscore"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$FI_AC1_n_correct_answers_Zscore))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_FI_AC1_n_correct_answers_Zscore'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### Covarying for SDS
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS, "+ SDS_online1_n_correct_cons_attempts_Zscore_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_SDS_online1_n_correct_cons_attempts_Zscore_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn, "+ SDS_online1_n_correct_cons_attempts_Zscore_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$SDS_online1_n_correct_cons_attempts_Zscore_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_SDS_online1_n_correct_cons_attempts_Zscore_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### Covarying for TMTA
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS, "+ TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn, "+ TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### Covarying for NM
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS, "+ Numeric_memory_online_maximum_digits_remembered_Zscore"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$Numeric_memory_online_maximum_digits_remembered_Zscore))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_Numeric_memory_online_maximum_digits_remembered_Zscore'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn, "+ Numeric_memory_online_maximum_digits_remembered_Zscore"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$Numeric_memory_online_maximum_digits_remembered_Zscore))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_Numeric_memory_online_maximum_digits_remembered_Zscore'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### Covarying for RT
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS, "+ Reaction_time_AC1_mean_time_taken_Zscore_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_Reaction_time_AC1_mean_time_taken_Zscore_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn, "+ Reaction_time_AC1_mean_time_taken_Zscore_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_Reaction_time_AC1_mean_time_taken_Zscore_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### Covarying for PM
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS, "+ Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn, "+ Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### Covarying for TMTB
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS, "+ TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn, "+ TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm"))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]) & !is.na(EU_ancs$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_covar_TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

## Other anc regressions
```{r}
sets <- list('pLI_0.9')
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')
```

### AAC regression
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

covars_AAC_ancs_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + SSA + AAC_PC1 + AAC_PC2 + AAC_PC3 + AAC_PC4 + AAC_PC5 + AAC_PC6 + AAC_PC7 + AAC_PC8 + AAC_PC9 + AAC_PC10' 

covars_AAC_ancs_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + SSA + AAC_PC1 + AAC_PC2 + AAC_PC3 + AAC_PC4 + AAC_PC5 + AAC_PC6 + AAC_PC7 + AAC_PC8 + AAC_PC9 + AAC_PC10' 


# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_AAC_ancs_NS))
      model <- lm(formula, data = AAC_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(AAC_ancs, !is.na(AAC_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'AAC_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_AAC_ancs_syn))
      model <- lm(formula, data = AAC_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(AAC_ancs, !is.na(AAC_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'AAC_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

### EAS regression
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

covars_EAS_ancs_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + SSA + EAS_PC1 + EAS_PC2 + EAS_PC3 + EAS_PC4 + EAS_PC5 + EAS_PC6 + EAS_PC7 + EAS_PC8 + EAS_PC9 + EAS_PC10' 

covars_EAS_ancs_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + SSA + EAS_PC1 + EAS_PC2 + EAS_PC3 + EAS_PC4 + EAS_PC5 + EAS_PC6 + EAS_PC7 + EAS_PC8 + EAS_PC9 + EAS_PC10' 


# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_EAS_ancs_NS))
      model <- lm(formula, data = EAS_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EAS_ancs, !is.na(EAS_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EAS_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_EAS_ancs_syn))
      model <- lm(formula, data = EAS_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EAS_ancs, !is.na(EAS_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EAS_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```


### LAT regression
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

covars_LAT_ancs_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + SSA + LAT_PC1 + LAT_PC2 + LAT_PC3 + LAT_PC4 + LAT_PC5 + LAT_PC6 + LAT_PC7 + LAT_PC8 + LAT_PC9 + LAT_PC10' 

covars_LAT_ancs_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + SSA + LAT_PC1 + LAT_PC2 + LAT_PC3 + LAT_PC4 + LAT_PC5 + LAT_PC6 + LAT_PC7 + LAT_PC8 + LAT_PC9 + LAT_PC10' 


# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_LAT_ancs_NS))
      model <- lm(formula, data = LAT_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(LAT_ancs, !is.na(LAT_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'LAT_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_LAT_ancs_syn))
      model <- lm(formula, data = LAT_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(LAT_ancs, !is.na(LAT_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'LAT_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

### SAS regression
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

covars_SAS_ancs_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + SSA + SAS_PC1 + SAS_PC2 + SAS_PC3 + SAS_PC4 + SAS_PC5 + SAS_PC6 + SAS_PC7 + SAS_PC8 + SAS_PC9 + SAS_PC10' 

covars_SAS_ancs_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + SSA + SAS_PC1 + SAS_PC2 + SAS_PC3 + SAS_PC4 + SAS_PC5 + SAS_PC6 + SAS_PC7 + SAS_PC8 + SAS_PC9 + SAS_PC10' 


# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_SAS_ancs_NS))
      model <- lm(formula, data = SAS_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(SAS_ancs, !is.na(SAS_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'SAS_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_SAS_ancs_syn))
      model <- lm(formula, data = SAS_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(SAS_ancs, !is.na(SAS_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'SAS_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

### SSA regression
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

covars_SSA_ancs_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + SSA + SSA_PC1 + SSA_PC2 + SSA_PC3 + SSA_PC4 + SSA_PC5 + SSA_PC6 + SSA_PC7 + SSA_PC8 + SSA_PC9 + SSA_PC10' 

covars_SSA_ancs_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + SSA + SSA_PC1 + SSA_PC2 + SSA_PC3 + SSA_PC4 + SSA_PC5 + SSA_PC6 + SSA_PC7 + SSA_PC8 + SSA_PC9 + SSA_PC10' 


# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_SSA_ancs_NS))
      model <- lm(formula, data = SSA_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(SSA_ancs, !is.na(SSA_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'SSA_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_SSA_ancs_syn))
      model <- lm(formula, data = SSA_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(SSA_ancs, !is.na(SSA_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'SSA_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

### ADM regression
```{r}
var_types<-list('PTV', 'REVEL75_Miss')

covars_ADM_ancs_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + SSA + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + PC11 + PC12 + PC13 + PC14 + PC15 + PC16 + PC17 + PC18 + PC19 + PC20 + PC21 + PC22 + PC23 + PC24 + PC25 + PC26 + PC27 + PC28 + PC29 + PC30 +PC31 + PC32 + PC33 + PC34' 

covars_ADM_ancs_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + SSA + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + PC11 + PC12 + PC13 + PC14 + PC15 + PC16 + PC17 + PC18 + PC19 + PC20 + PC21 + PC22 + PC23 + PC24 + PC25 + PC26 + PC27 + PC28 + PC29 + PC30 +PC31 + PC32 + PC33 + PC34' 


# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_ADM_ancs_NS))
      model <- lm(formula, data = ADM_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(ADM_ancs, !is.na(ADM_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'ADM_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


var_types<-list('synonymous')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_ADM_ancs_syn))
      model <- lm(formula, data = ADM_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(ADM_ancs, !is.na(ADM_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'ADM_ancs'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```


# Save out results table 
```{r}
write.csv(results, 'results_g_regressions_AC5orless_diagnosesrm_incSE_crossancs_Feb25.csv')
system('dx upload results_g_regressions_AC5orless_diagnosesrm_incSE_crossancs_Feb25.csv --destination ./Variant_counts/Gene_sets/')
```
