---
title: "18_single_gene_regressions"
output:
  pdf_document: default
  html_document: default
date: "2024-10-01"
---

# Single gene regressions
This script reads in phenotypic data and tables of counts, and then runs regressions on these data. This script was run on a mem2_ssd2_v2_x16 RStudio instance, and took 3h30mins to run costing £2.20. 


# Load libraries
```{r}
library(data.table)
library(dplyr)
library(purrr)
install.packages('broom')
library(broom)

```

# Prep data
## Read in cognitive phenotypes
```{r}
system('dx download ./phenotype_cohorts/processed_cognition/processed_cog_tests_inc_g_Feb24.csv')
phen=read.csv('processed_cog_tests_inc_g_Feb24.csv', header=T)

```

## Ancestries
```{r}
system('dx download ./genetic_ancestries/app15175_ids/full_dataset_inferred_ancestries.csv')
ancs_wrongid=read.csv('full_dataset_inferred_ancestries.csv', header=T)
system('dx download ./genetic_ancestries/app15175_ids/PCAs_*')
EUR_PCs_wrongid=read.table('PCAs_EUR_only.tsv', header=T)
AAC_PCs_wrongid=read.table('PCAs_AAC_only.tsv', header=T)
EAS_PCs_wrongid=read.table('PCAs_EAS_only.tsv', header=T)
LAT_PCs_wrongid=read.table('PCAs_LAT_only.tsv', header=T)
SAS_PCs_wrongid=read.table('PCAs_SAS_only.tsv', header=T)
SSA_PCs_wrongid=read.table('PCAs_SSA_only.tsv', header=T)
ADM_PCs_wrongid=read.table('PCAs_ADM_only.tsv', header=T)

system('dx download ./phenotype_cohorts/link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt')
link_ids=read.table('link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt', header=T)
ancs=merge(ancs_wrongid, link_ids, by.x='FID', by.y='app15175')
ancs=merge(ancs, EUR_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, AAC_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, EAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, LAT_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SSA_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, ADM_PCs_wrongid, by.x ='FID', by.y='s', all=T)

# Write out now have correct IDs
write.csv(ancs, 'ancestries_PCs.csv')
system('dx upload ancestries_PCs.csv ./genetic_ancestries/')

# Merge 
phen=merge(phen, ancs, by.x = 'eid', by.y='eid_13310')

```

## QC
Filter to those passing the final QC stage (sex imputation)
```{r}
system('dx download ./WES_QC/discordant_sex_ids.csv')
ids_to_rm=read.csv('discordant_sex_ids.csv')
ids_to_rm$remove=T

phen=merge(phen, ids_to_rm, by.x = 'eid', by.y='x', all=T)
phen=subset(phen, is.na(phen$remove)) 

```

## Covariates
Read in covariates (and prep for use). These phenotypes were selected in cohort browser and written out using the Table Exporter tool. 
```{r}
system('dx download ./phenotype_cohorts/covars_participant.csv')
covars=read.csv('covars_participant.csv', header=T)

covars$seq_batch<-covars$p32050
covars$AC_attended<-covars$p54_i0
covars$sex_1male[covars$p31 == 'Male']<-1
covars$sex_1male[covars$p31 == 'Female']<-2
covars$standardised_age <- (covars$p21022 - mean(covars$p21022, na.rm = TRUE)) / sd(covars$p21022, na.rm = TRUE)
covars$standardised_age_squared<-covars$standardised_age*covars$standardised_age
covars$sex_standardised_age<-covars$sex_1male * covars$standardised_age
covars$sex_standardised_age_squared<-covars$sex_1male * covars$standardised_age_squared


phen=merge(phen, covars, by.x='eid', by.y='eid')


```

## Diagnoses 
These phenotypes were selected in cohort browser and written out using the Table Exporter tool. 
```{r}
system('dx download ./phenotypes_cog_diagnoses_participation_death.csv')
diagnoses=read.csv('phenotypes_cog_diagnoses_participation_death.csv', header=T)

# Make subset of people diagnosed with SZ - to run these regressions in if powered? 
sz_diagnosed=subset(diagnoses, diagnoses$SZ==T) 
sz_diagnosed_eur_only=subset(sz_diagnosed, sz_diagnosed$EUR>0.8)

# Remove people with SZ, ID and/or ASD 
diagnoses=subset(diagnoses, diagnoses$SZ==F & diagnoses$ASD==F & diagnoses$intellectual_disability==F) 
diagnoses$no_diagnoses<-T
diagnoses=diagnoses[,c(3,174)]

phen=merge(phen, diagnoses, by='eid') 
```


# AC5orless -  single gene regression
Each type of variant is run separately below, as this process takes a while and it's good to see the progess, and know saving out has worked after each variant. 
## Prepare covars
```{r}
AC = '5orless'

system('dx download ./Variant_counts/Gene_sets/Feb25/exome_wide_synonymous_AC5orless_counts.csv')
a=read.csv('exome_wide_synonymous_AC5orless_counts.csv')
phen=merge(phen, a, by.x='eid', by.y='Sample_ID')

# List covars
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

```


## PTVs
```{r}
# Read in 
system('dx download ./Variant_counts/Gene_sets/single_gene/single_gene_tests_PTV_AC5orless_counts.csv')
ptv_single_gene=read.csv('single_gene_tests_PTV_AC5orless_counts.csv')
all_ancs=merge(phen, ptv_single_gene, by.x='eid', by.y='Sample_ID')
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

# Set up gene list 
genes=ptv_single_gene[,-2]
genes=genes[,-1]
sets <- as.list(colnames(genes))
var<-'PTV'
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

# Set up empty results table
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)


# Loop through each gene
for (dep_var in dependent_var){
  for (set in sets) {
      independent_var <- paste0(set)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE=SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
  }
}

write.csv(results, 'PTV_EUR_anc_AC5orless_single_gene_burden.csv')
system('dx upload PTV_EUR_anc_AC5orless_single_gene_burden.csv --destination ./Variant_counts/Gene_sets/single_gene/')
```
## REVEL >0.75 missense
```{r}
system('dx download ./Variant_counts/Gene_sets/single_gene/single_gene_tests_REVEL75_Miss_AC5orless_counts.csv')

#Clear big ptv file first
rm(ptv_single_gene)
gc()

# Read in
miss_single_gene=read.csv('single_gene_tests_REVEL75_Miss_AC5orless_counts.csv')
all_ancs=merge(phen, miss_single_gene, by.x='eid', by.y='Sample_ID')
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

# Set up gene list 
genes=miss_single_gene[,-2]
genes=genes[,-1]
sets <- as.list(colnames(genes))
var<-'REVEL75_Miss'
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

# Set up empty results table
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)


# Loop through each gene
for (dep_var in dependent_var){
  for (set in sets) {
      independent_var <- paste0(set)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE=SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
  }
}

write.csv(results, 'REVEL75_Miss_EUR_anc_AC5orless_single_gene_burden.csv')
system('dx upload REVEL75_Miss_EUR_anc_AC5orless_single_gene_burden.csv --destination ./Variant_counts/Gene_sets/single_gene/')
```


## REVEL miss + PTVS
```{r}
system('dx download ./Variant_counts/Gene_sets/single_gene/single_gene_tests_PTVplusMiss_AC5orless_counts.csv')
both_single_gene=read.csv('single_gene_tests_PTVplusMiss_AC5orless_counts.csv')
all_ancs=merge(phen, both_single_gene, by.x='eid', by.y='Sample_ID')
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

# Set up gene list 
genes=both_single_gene[,-2]
genes=genes[,-1]
sets <- as.list(colnames(genes))
var<-'PTVs_and_miss'
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

# Set up empty results table
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)


# Loop through each gene
for (dep_var in dependent_var){
  for (set in sets) {
      independent_var <- paste0(set)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE=SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
  }
}

write.csv(results, 'PTVs_and_Miss_EUR_anc_AC5orless_single_gene_burden.csv')
system('dx upload PTVs_and_Miss_EUR_anc_AC5orless_single_gene_burden.csv --destination ./Variant_counts/Gene_sets/single_gene/')
```
