---
title: "19_combined_analysis"
output: html_document
date: "2024-10-01"
---

# Combined analysis
This script runs a combined analysis of RCVs, CNVs and PRS. This script ran for 40 minutes on a mem1_ssd1_v2_x8 RStudio instance, costing £0.40. 
# Load libraries
```{r}
library(data.table)
library(dplyr)
library(purrr)
install.packages('broom')
library(broom)
library(ggplot2)
```

# Prep data
## Read in cognitive phenotypes
```{r}
system('dx download ./phenotype_cohorts/processed_cognition/processed_cog_tests_inc_g_Feb24.csv')
phen=read.csv('processed_cog_tests_inc_g_Feb24.csv', header=T)

```

## Ancestries
```{r}
system('dx download ./genetic_ancestries/app15175_ids/full_dataset_inferred_ancestries.csv')
ancs_wrongid=read.csv('full_dataset_inferred_ancestries.csv', header=T)
system('dx download ./genetic_ancestries/app15175_ids/PCAs_*')
EUR_PCs_wrongid=read.table('PCAs_EUR_only.tsv', header=T)
AAC_PCs_wrongid=read.table('PCAs_AAC_only.tsv', header=T)
EAS_PCs_wrongid=read.table('PCAs_EAS_only.tsv', header=T)
LAT_PCs_wrongid=read.table('PCAs_LAT_only.tsv', header=T)
SAS_PCs_wrongid=read.table('PCAs_SAS_only.tsv', header=T)
SSA_PCs_wrongid=read.table('PCAs_SSA_only.tsv', header=T)
ADM_PCs_wrongid=read.table('PCAs_ADM_only.tsv', header=T)

system('dx download ./phenotype_cohorts/link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt')
link_ids=read.table('link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt', header=T)
ancs=merge(ancs_wrongid, link_ids, by.x='FID', by.y='app15175')
ancs=merge(ancs, EUR_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, AAC_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, EAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, LAT_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SSA_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, ADM_PCs_wrongid, by.x ='FID', by.y='s', all=T)

# Write out now have correct IDs
write.csv(ancs, 'ancestries_PCs.csv')
system('dx upload ancestries_PCs.csv ./genetic_ancestries/')

# Merge 
phen=merge(phen, ancs, by.x = 'eid', by.y='eid_13310')

```

## QC
Filter to those passing the final QC stage (sex imputation)
```{r}
system('dx download ./WES_QC/discordant_sex_ids.csv')
ids_to_rm=read.csv('discordant_sex_ids.csv')
ids_to_rm$remove=T

phen=merge(phen, ids_to_rm, by.x = 'eid', by.y='x', all=T)
phen=subset(phen, is.na(phen$remove)) 

```

## Covariates
Read in covariates (and prep for use). These phenotypes were selected in cohort browser and written out using the Table Exporter tool. 
```{r}
system('dx download ./phenotype_cohorts/covars_participant.csv')
covars=read.csv('covars_participant.csv', header=T)

covars$seq_batch<-covars$p32050
covars$AC_attended<-covars$p54_i0
covars$sex_1male[covars$p31 == 'Male']<-1
covars$sex_1male[covars$p31 == 'Female']<-2
covars$standardised_age <- (covars$p21022 - mean(covars$p21022, na.rm = TRUE)) / sd(covars$p21022, na.rm = TRUE)
covars$standardised_age_squared<-covars$standardised_age*covars$standardised_age
covars$sex_standardised_age<-covars$sex_1male * covars$standardised_age
covars$sex_standardised_age_squared<-covars$sex_1male * covars$standardised_age_squared


phen=merge(phen, covars, by.x='eid', by.y='eid')


```

## Diagnoses 
These phenotypes were selected in cohort browser and written out using the Table Exporter tool. 
```{r}
system('dx download ./phenotypes_cog_diagnoses_participation_death.csv')
diagnoses=read.csv('phenotypes_cog_diagnoses_participation_death.csv', header=T)

# Make subset of people diagnosed with SZ - to run these regressions in if powered? 
sz_diagnosed=subset(diagnoses, diagnoses$SZ==T) 
sz_diagnosed_eur_only=subset(sz_diagnosed, sz_diagnosed$EUR>0.8)

# Remove people with SZ, ID and/or ASD 
diagnoses=subset(diagnoses, diagnoses$SZ==F & diagnoses$ASD==F & diagnoses$intellectual_disability==F) 
diagnoses$no_diagnoses<-T
diagnoses=diagnoses[,c(3,174)]

phen=merge(phen, diagnoses, by='eid') 
```



## RCVs 
Here were looking at RCVs in LoFi genes 
```{r}
AC='5orless'
system('dx download ./Variant_counts/Gene_sets/pLI*_AC5orless_*.csv --overwrite')
file_list <- list.files(path = "./", pattern = "*_AC5orless_counts.csv", full.names = TRUE)
# Read, drop first column (V1), and merge each file
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames

```



## CNVs 
```{r}
system('dx download ./NDD_CNVs/Pathogenic_CNVs.csv')
CNVs=read.csv('Pathogenic_CNVs.csv')
CNVs$has_CNV_data<-T
CNVs %>% count(filter_good_CNVs)
CNVs$carries_HQ_CNV<-F
CNVs$carries_HQ_CNV[CNVs$filter_good_CNVs=='Selected']<-T


CNVs$SZ_enriched_CNV_carrier <- TRUE
CNVs$SZ_enriched_CNV_carrier[CNVs$carries_HQ_CNV==F]<-NA
CNVs %>% count(SZ_enriched_CNV_carrier)

CNVs %>% count(Pathogenic_checked)
CNVs$SZ_associated_CNV_carrier<-FALSE
CNVs$SZ_associated_CNV_carrier[CNVs$Pathogenic_checked=='1q21.1del'|
                               CNVs$Pathogenic_checked=='1q21.1dup'|
                               CNVs$Pathogenic_checked=='NRXN1del'|
                               CNVs$Pathogenic_checked=='3q29del'|
                               CNVs$Pathogenic_checked=='WBS_dup'|
                               CNVs$Pathogenic_checked=='PWS_dup'|
                               CNVs$Pathogenic_checked=='15q11.2del'|
                               CNVs$Pathogenic_checked=='15q13.3del'|
                               CNVs$Pathogenic_checked=='16p13.11dup'|
                               CNVs$Pathogenic_checked=='16p12.1del'|
                               CNVs$Pathogenic_checked=='16p11.2distal_del'|
                               CNVs$Pathogenic_checked=='16p11.2dup'|
                               CNVs$Pathogenic_checked=='22q11.2del']<-TRUE
CNVs$SZ_associated_CNV_carrier[CNVs$carries_HQ_CNV==F]<-NA
CNVs %>% count(SZ_associated_CNV_carrier)


CNVs$SZ_associated_CNV_carrier_no15q<-FALSE
CNVs$SZ_associated_CNV_carrier_no15q[CNVs$Pathogenic_checked=='1q21.1del'|
                               CNVs$Pathogenic_checked=='1q21.1dup'|
                               CNVs$Pathogenic_checked=='NRXN1del'|
                               CNVs$Pathogenic_checked=='3q29del'|
                               CNVs$Pathogenic_checked=='WBS_dup'|
                               CNVs$Pathogenic_checked=='PWS_dup'|
                               CNVs$Pathogenic_checked=='15q13.3del'|
                               CNVs$Pathogenic_checked=='16p13.11dup'|
                               CNVs$Pathogenic_checked=='16p12.1del'|
                               CNVs$Pathogenic_checked=='16p11.2distal_del'|
                               CNVs$Pathogenic_checked=='16p11.2dup'|
                               CNVs$Pathogenic_checked=='22q11.2del']<-TRUE
CNVs$SZ_associated_CNV_carrier_no15q[CNVs$carries_HQ_CNV==F]<-NA
CNVs %>% count(SZ_associated_CNV_carrier_no15q)


CNVs$CNV_carrier_15q_only<-FALSE
CNVs$CNV_carrier_15q_only[CNVs$Pathogenic_checked=='15q11.2del']<-TRUE
CNVs$CNV_carrier_15q_only[CNVs$carries_HQ_CNV==F]<-NA
CNVs %>% count(CNV_carrier_15q_only)

CNVs_simple = CNVs[,c('f.eid','has_CNV_data', 
                      'carries_HQ_CNV', 'SZ_enriched_CNV_carrier', 
                      'SZ_associated_CNV_carrier', 'CNV_carrier_15q_only')]

# Look at the people carrying >1 cnv
duplicates <- CNVs_simple %>%
  group_by(f.eid) %>%
  filter(n() > 1)
duplicates %>% count(f.eid) 

# Make df with 1 row per person (as we're just interested in whether they're a carrier, rather than how many CNVs they carry)
CNVs_summary <- CNVs_simple %>%
  group_by(f.eid) %>%
  summarise(across(everything(), any))  # Keeps TRUE if it appears in any row for that f.eid

# Read in link ids file 
system('dx download ./phenotype_cohorts/link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt')
link_ids=read.table('link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt', header=T)
CNVs_summary=merge(CNVs_summary, link_ids, by.x='f.eid', by.y='app14421')
CNVs_summary=CNVs_summary[,c(9,2:6)]

combined = merge(counts, CNVs_summary, by.x = 'Sample_ID', by.y = 'eid_13310', all.x=T) 
combined$SZ_associated_CNV_carrier[is.na(combined$SZ_associated_CNV_carrier)]<-F
combined$SZ_associated_CNV_carrier[combined$carries_HQ_CNV==F]<-NA
combined %>% count(SZ_associated_CNV_carrier) # 2,986

combined$SZ_enriched_CNV_carrier[is.na(combined$SZ_enriched_CNV_carrier)]<-F
combined$SZ_enriched_CNV_carrier[combined$carries_HQ_CNV==F]<-NA
combined %>% count(SZ_enriched_CNV_carrier) # 4,282

combined$CNV_carrier_15q_only[is.na(combined$CNV_carrier_15q_only)]<-F
combined$CNV_carrier_15q_only[combined$carries_HQ_CNV==F]<-NA
combined %>% count(CNV_carrier_15q_only) # 1395

```

## PRS
```{r}
system('dx download ./SZ_PRS/*')
UKB_derived_PRS=read.csv('data_participant.csv')
UKB_derived_PRS$UKB_standard_SZ_PRS<- UKB_derived_PRS$p26275
UKB_derived_PRS=UKB_derived_PRS[,c(1,5)]


SL_derived_PRS=read.table('ukbb.multiancestry.SZ.prsice.all_score', header = T)
SL_derived_PRS$SL_SZ_PRS <- SL_derived_PRS$Pt_0.05
SZ_PRS<- SL_derived_PRS[,c('FID', 'SL_SZ_PRS')] # 487409

# Check correlations?
SZ_PRS=merge(SZ_PRS, link_ids, by.x='FID', by.y='app15175') #
SZ_PRS=SZ_PRS[,c(5,2)]
SZ_PRS<-merge(SZ_PRS, UKB_derived_PRS, by.x='eid_13310', by.y = 'eid', all=T)
a<- SZ_PRS[,c(2,3)]
correlations <- cor(a, use="pairwise.complete.obs") # 0.4

combined = merge(combined, SZ_PRS, by.x = 'Sample_ID', by.y = 'eid_13310') 

# Remove people with low quality data for any variant type 
# 399,877 people with HQ sequencing data
combined_HQ=subset(combined, !is.na(combined$SZ_enriched_CNV_carrier)) 
combined_HQ=subset(combined_HQ, !is.na(combined_HQ$UKB_standard_SZ_PRS)) 



```


## Merge all together 
```{r}
all_ancs=merge(phen, combined_HQ, by.x='eid', by.y='Sample_ID')
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8)
EU_ancs$ST_SL_SZ_PRS<- (EU_ancs$SL_SZ_PRS - mean(EU_ancs$SL_SZ_PRS, na.rm = TRUE)) / sd(EU_ancs$SL_SZ_PRS, na.rm = TRUE)

array_batch=read.csv('array_batch_participant.csv')
array_batch %>% count(array_batch$p22000)
array_batch=subset(array_batch, array_batch$p22000!='')
array_batch$genotype_batch<-array_batch$p22000
array_batch=array_batch[,c(1,3)]
EU_ancs=merge(EU_ancs, array_batch, by='eid')


EU_ancs_w_g=subset(EU_ancs, !is.na(EU_ancs$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm)) # 75,111
backup=EU_ancs_w_g
```

# Run regressions!!
## Prep covars
```{r}
system('dx download ./Variant_counts/Gene_sets/Feb25/exome_wide_synonymous_AC5orless_counts.csv')
a=read.csv('exome_wide_synonymous_AC5orless_counts.csv')
EU_ancs_w_g=merge(EU_ancs_w_g, a, by.x='eid', by.y='Sample_ID')

covars = 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch'
```

## Multivariable
``` {r}
# full model
multivar_model <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       PTV_AC5orless_pLI_0.9_total + 
                       REVEL75_Miss_AC5orless_pLI_0.9_total + 
                       SZ_enriched_CNV_carrier +
                       ST_SL_SZ_PRS +
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
tidy_model <- tidy(multivar_model)
CIs<- confint(multivar_model)

adj_r2_multivar <- summary(multivar_model)$adj.r.squared # 0.17689






# minus each predictor to get r2 
multivar_model_minusPTVs <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       REVEL75_Miss_AC5orless_pLI_0.9_total + 
                       SZ_enriched_CNV_carrier +
                       ST_SL_SZ_PRS +
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
summary(multivar_model)$adj.r.squared - summary(multivar_model_minusPTVs)$adj.r.squared





multivar_model_minusmiss <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       PTV_AC5orless_pLI_0.9_total + 
                       SZ_enriched_CNV_carrier +
                       ST_SL_SZ_PRS +
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
summary(multivar_model)$adj.r.squared - summary(multivar_model_minusmiss)$adj.r.squared






multivar_model_minusCNV <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       PTV_AC5orless_pLI_0.9_total + 
                       REVEL75_Miss_AC5orless_pLI_0.9_total + 
                       ST_SL_SZ_PRS +
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
summary(multivar_model)$adj.r.squared - summary(multivar_model_minusCNV)$adj.r.squared



multivar_model_minusPRS <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       PTV_AC5orless_pLI_0.9_total + 
                       REVEL75_Miss_AC5orless_pLI_0.9_total + 
                       SZ_enriched_CNV_carrier +
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
summary(multivar_model)$adj.r.squared - summary(multivar_model_minusPRS)$adj.r.squared


multivar_model_minusAllgeneticfactors <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
summary(multivar_model)$adj.r.squared - summary(multivar_model_minusAllgeneticfactors)$adj.r.squared



```

# Univariable
## Get r2 for just covar model 
```{r}
just_covar_model <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
adj_r2_justcovar <- summary(just_covar_model)$adj.r.squared # 0.1740099

```

## PTVs
``` {r}
univar_model <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       PTV_AC5orless_pLI_0.9_total + 
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
tidy_model <- tidy(univar_model)
CIs<- confint(univar_model)

summary(univar_model)$adj.r.squared - adj_r2_justcovar
```


## Miss
``` {r}
univar_model <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       REVEL75_Miss_AC5orless_pLI_0.9_total + 
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
tidy_model <- tidy(univar_model)
CIs<- confint(univar_model)

summary(univar_model)$adj.r.squared - adj_r2_justcovar
```

## PRS
``` {r}
univar_model <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       ST_SL_SZ_PRS + 
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
tidy_model <- tidy(univar_model)
CIs<- confint(univar_model)

summary(univar_model)$adj.r.squared - adj_r2_justcovar
```

## SZ enriched CNVs
``` {r}
univar_model <- lm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ 
                       SZ_enriched_CNV_carrier + 
                       seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + 
                       sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + 
                       AAC + EAS + LAT + SAS + 
                       EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + 
                       EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch, 
                     data = EU_ancs_w_g) 
tidy_model <- tidy(univar_model)
CIs<- confint(univar_model)

summary(univar_model)$adj.r.squared - adj_r2_justcovar
```
