---
title: "review_response"
output: html_document
date: "2025-10-01"
---
# Load libraries
```{r}
library(dplyr)
library(ggplot2)
library(data.table)
library(purrr)
install.packages('broom')
library(broom)
library(purrr)
library(tidyr)
```
# g in people with an NPD 
## Read in
```{r}
### Read in
system('dx download ./phenotype_cohorts/processed_cognition/processed_cog_tests_inc_g_Feb24.csv')
phen=read.csv('processed_cog_tests_inc_g_Feb24.csv', header=T)
## Ancestries
system('dx download ./genetic_ancestries/app15175_ids/full_dataset_inferred_ancestries.csv')
ancs_wrongid=read.csv('full_dataset_inferred_ancestries.csv', header=T)
system('dx download ./genetic_ancestries/app15175_ids/PCAs_*')
EUR_PCs_wrongid=read.table('PCAs_EUR_only.tsv', header=T)
AAC_PCs_wrongid=read.table('PCAs_AAC_only.tsv', header=T)
EAS_PCs_wrongid=read.table('PCAs_EAS_only.tsv', header=T)
LAT_PCs_wrongid=read.table('PCAs_LAT_only.tsv', header=T)
SAS_PCs_wrongid=read.table('PCAs_SAS_only.tsv', header=T)
SSA_PCs_wrongid=read.table('PCAs_SSA_only.tsv', header=T)
ADM_PCs_wrongid=read.table('PCAs_ADM_only.tsv', header=T)

system('dx download ./phenotype_cohorts/link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt')
link_ids=read.table('link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt', header=T)
ancs=merge(ancs_wrongid, link_ids, by.x='FID', by.y='app15175')
ancs=merge(ancs, EUR_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, AAC_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, EAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, LAT_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SSA_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, ADM_PCs_wrongid, by.x ='FID', by.y='s', all=T)

# Merge 
phen=merge(phen, ancs, by.x = 'eid', by.y='eid_13310')

## QC
system('dx download ./WES_QC/discordant_sex_ids.csv')
ids_to_rm=read.csv('discordant_sex_ids.csv')
ids_to_rm$remove=T
phen=merge(phen, ids_to_rm, by.x = 'eid', by.y='x', all=T)
phen=subset(phen, is.na(phen$remove)) 


## Covariates
system('dx download ./phenotype_cohorts/covars_participant.csv')
covars=read.csv('covars_participant.csv', header=T)
covars$seq_batch<-covars$p32050
covars$AC_attended<-covars$p54_i0
covars$sex_1male[covars$p31 == 'Male']<-1
covars$sex_1male[covars$p31 == 'Female']<-2
covars$standardised_age <- (covars$p21022 - mean(covars$p21022, na.rm = TRUE)) / sd(covars$p21022, na.rm = TRUE)
covars$standardised_age_squared<-covars$standardised_age*covars$standardised_age
covars$sex_standardised_age<-covars$sex_1male * covars$standardised_age
covars$sex_standardised_age_squared<-covars$sex_1male * covars$standardised_age_squared
phen=merge(phen, covars, by.x='eid', by.y='eid')


## Diagnoses
system('dx download ./phenotypes_cog_diagnoses_participation_death.csv')
diagnoses=read.csv('phenotypes_cog_diagnoses_participation_death.csv', header=T)

# Remove people with SZ, ID and/or ASD 
diagnoses$NPD[diagnoses$SZ==F & diagnoses$ASD==F & diagnoses$intellectual_disability==F] <- F
diagnoses$NPD[diagnoses$SZ==T | diagnoses$ASD==T | diagnoses$intellectual_disability==T] <- T
diagnoses %>% count(NPD) #1,528

diagnoses=diagnoses[,c(3,174)]

phen=merge(phen, diagnoses, by='eid') 
```

## Plot 
```{r}
with_g = subset(phen, !is.na(phen$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm)) #76,888
with_g %>% count(NPD)
g_SZ=subset(with_g, with_g$NPD==T)
mean(g_SZ$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm)
sd(g_SZ$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm)/sqrt(nrow(g_SZ))



g_no_SZ=subset(with_g, with_g$NPD==F)
mean(g_no_SZ$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm)
sd(g_no_SZ$g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm)/sqrt(nrow(g_no_SZ))

# Plot mean and se
summary_df <- with_g %>%
  group_by(NPD) %>%
  summarise(
    mean_g = mean(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm, na.rm = TRUE),
    se_g = sd(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm, na.rm = TRUE) / sqrt(n())
  )



ggplot(summary_df, aes(x = factor(NPD), y = mean_g, color = factor(NPD))) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = mean_g - se_g, ymax = mean_g + se_g), width = 0.2) +
  scale_y_continuous(limits = c(-0.36, 0.05)) +
  labs(x = "Group",
       y =  expression(paste("Mean ", italic("g")))) +
  theme_minimal(base_size = 14)+
  scale_x_discrete(labels = c("FALSE" = "No diagnosis of \n SZ, ASD or ID",
                            "TRUE"  = "Diagnosis of SZ, \n ASD, and/or ID"))+
  theme(legend.position = "none")



```

## Statistically compare
```{r}
# All ancestries
model_all_anc <- glm(g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm ~ NPD + 
                       AC_attended + sex_1male + standardised_age + 
                       standardised_age_squared + sex_standardised_age + 
                       sex_standardised_age_squared + AAC + EAS + LAT + SAS + 
                       SSA + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + 
                       PC9 + PC10 + PC11 + PC12 + PC13 + PC14 + PC15 + PC16 + 
                       PC17 + PC18 + PC19 + PC20 + PC21 + PC22 + PC23 + PC24 + 
                       PC25 + PC26 + PC27 + PC28 + PC29 + PC30 + PC31 + PC32 + 
                       PC33 + PC34,
                     data = with_g)
a= tidy(model_all_anc)


```

# Explaining variance in FI 
```{r}
with_g_with_FI_AC1=subset(with_g, !is.na(with_g$FI_AC1_n_correct_answers_Zscore))
m1 <- lm(FI_AC1_n_correct_answers_Zscore ~ AC_attended + sex_1male + standardised_age + 
                       standardised_age_squared + sex_standardised_age + 
                       sex_standardised_age_squared, data = with_g)
summary(m1)

m2<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm + AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m2)
summary(m2)$adj.r.squared - summary(m1)$adj.r.squared

m3<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          Numeric_memory_AC1_maximum_digits_remembered_Zscore + AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m3)
summary(m3)$adj.r.squared - summary(m1)$adj.r.squared

m4<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm + AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m4)
summary(m4)$adj.r.squared - summary(m1)$adj.r.squared


m5<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          Reaction_time_AC1_mean_time_taken_Zscore_outliersrm + AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m5)
summary(m5)$adj.r.squared - summary(m1)$adj.r.squared


m6<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm + AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m6)
summary(m6)$adj.r.squared - summary(m1)$adj.r.squared

m7<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm + AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m7)
summary(m7)$adj.r.squared - summary(m1)$adj.r.squared


m8<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          SDS_online1_n_correct_cons_attempts_Zscore_outliersrm + AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m8)
summary(m8)$adj.r.squared - summary(m1)$adj.r.squared

m9<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          prospective_mem_AC1 + AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m9)
summary(m9)$adj.r.squared - summary(m1)$adj.r.squared


m_all<- lm(FI_AC1_n_correct_answers_Zscore ~ 
          prospective_mem_AC1 + 
          SDS_online1_n_correct_cons_attempts_Zscore_outliersrm +
            TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm +
            TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm +
            Reaction_time_AC1_mean_time_taken_Zscore_outliersrm +
            Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm +
            Numeric_memory_AC1_maximum_digits_remembered_Zscore +
            g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm + 
          AC_attended +
          sex_1male + standardised_age + 
          standardised_age_squared + sex_standardised_age + 
          sex_standardised_age_squared, data = with_g)
summary(m_all)
summary(m_all)$adj.r.squared - summary(m1)$adj.r.squared    

with_g$TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm
```


# SMR expression separated gene sets
Run script 14 and 16 for these sets and then plot. 
## Read data for gene set counts
```{r}
# Set variables 
AC='5orless' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('PTV', 'REVEL75_Miss', 'synonymous')

# Download data
# PTVs
var<-'PTV'
input_filename <- paste0('./Variant_counts/transposed_counts/AC', AC, '/AC', AC, '_', var,'/transposed*.tsv')
download_command <- paste('dx download ', shQuote(input_filename))
system(download_command)

# Pathogenic missense
var<-'REVEL75_Miss'
input_filename <- paste0('./Variant_counts/transposed_counts/AC', AC, '/AC', AC, '_', var,'/transposed*.tsv')
download_command <- paste('dx download ', shQuote(input_filename))
system(download_command)

# Synonymous variants
var<-'synonymous'
input_filename <- paste0('./Variant_counts/transposed_counts/AC', AC, '/AC', AC, '_', var,'/transposed*.tsv')
download_command <- paste('dx download ', shQuote(input_filename))
system(download_command)


# Gene set annotations
system('dx download Variant_counts/Gene_sets/GWAS_sets_for_RAP_RR_Nov25.csv')
sets=read.csv('GWAS_sets_for_RAP_RR_Nov25.csv', header=T)
```

## Count across SMR sets
### Positive
```{r}
setname='SMR_prioritised_positive_expression'
set=subset(sets, sets$positive_expression==T) # 24 genes
list=set$gene_id

for (var in variant_classes){
  print(paste0('Running variant class ', var))
  for (chr in chromosomes) {
    # Process chr1 differently (as it's in 2 halves)
    if (chr==1) {
        # Read in 
        first_half <- paste0('transposed_chr_1_first_half_', var, '_AC', AC,'_gene_counts.tsv')
        first_half <- fread(first_half, header=TRUE, stringsAsFactors=FALSE)
        second_half <- paste0('transposed_chr_1_second_half_', var, '_AC', AC,'_gene_counts.tsv')
        second_half <- fread(second_half, header=TRUE, stringsAsFactors=FALSE)

        # Merge to one table and add up the gene split across the two halves
        chr1 <- merge(first_half, second_half, by = "Sample_ID")
        chr1$ENSG00000198758 <- chr1$ENSG00000198758.x + chr1$ENSG00000198758.y
        chr1 <- chr1 %>% select(-ENSG00000198758.x, -ENSG00000198758.y)
        cat('n genes in chr 1:', length(chr1) - 1, '\n')
        
        # Filter to your gene set 
        chr1 <- chr1 %>% select(c("Sample_ID", any_of(list)))
        cat('n genes in chr 1 in set:', length(chr1) - 1, '\n')
        # Only proceed if there are genes in the filtered set
        if (ncol(chr1) > 1) {
            chr1[[paste0("chr_1_", var, "_AC", AC, "_total")]] <- rowSums(chr1[,-1])
            geneset <- select(chr1, c("Sample_ID", paste0("chr_1_", var, "_AC", AC, "_total")))
        } else {
            cat("No genes from the gene set found in chr 1.\n")
        }
        rm(chr1)
        rm(first_half)
        rm(second_half)
        gc()
    }
    else {
        # All other chromosomes
        filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC,'_gene_counts.tsv')
        input <- fread(filename, header=TRUE, stringsAsFactors=FALSE)
        print(paste0('Processing chr ', chr))
        cat('n genes in chr:', length(input) - 1, '\n')
        # Filter the columns to keep only the genes present in the gene list
        chr_set <- input %>% select(c("Sample_ID", any_of(list)))
        cat('n genes in set in chr:' , length(chr_set) -1, '\n')
        # Add up variants in genes in set 
        if (ncol(chr_set) > 1) {
            chr_set[[paste0("chr_", chr, "_", var, "_AC", AC, "_total")]] <- rowSums(chr_set[,-1])
            chr_set <- select(chr_set, c("Sample_ID", paste0("chr_", chr, "_", var, "_AC", AC, "_total")))
            
            # Merge to big table and print length (to show increasing in size!!)
            geneset <- merge(geneset, chr_set, by='Sample_ID', all=TRUE)  
            # 'all=TRUE' to retain all rows in case one file is missing data
            cat('n vars in set in chr:', sum(geneset[[paste0("chr_", chr, "_", var, "_AC", AC, "_total")]]), '\n')
        } else {
            cat("No genes from the gene set found in chr", chr, ".\n")
        }
        
        # Clear mem
        rm(input)
        rm(chr_set)
        gc()}
    }

  # Add up across all chromosomes 
  geneset[[paste0(var,"_AC", AC, "_", setname, "_total")]] <- rowSums(geneset[,-1])
  geneset <- select(geneset, c("Sample_ID", paste0(var,"_AC", AC, "_", setname, "_total")))
  print(sum(geneset[[2]]))

  # Save out 
  output <- paste0(setname, '_', var, '_AC', AC, '_counts.csv')
  write.csv(geneset, output)


  upload_command <- paste('dx upload ', output, '--destination ./Variant_counts/Gene_sets/')
  system(upload_command)
}

```

### Negative
```{r}
setname='SMR_prioritised_negative_expression'
set=subset(sets, sets$negative_expression==T) #20 genes
list=set$gene_id

for (var in variant_classes){
  print(paste0('Running variant class ', var))
  for (chr in chromosomes) {
    # Process chr1 differently (as it's in 2 halves)
    if (chr==1) {
        # Read in 
        first_half <- paste0('transposed_chr_1_first_half_', var, '_AC', AC,'_gene_counts.tsv')
        first_half <- fread(first_half, header=TRUE, stringsAsFactors=FALSE)
        second_half <- paste0('transposed_chr_1_second_half_', var, '_AC', AC,'_gene_counts.tsv')
        second_half <- fread(second_half, header=TRUE, stringsAsFactors=FALSE)

        # Merge to one table and add up the gene split across the two halves
        chr1 <- merge(first_half, second_half, by = "Sample_ID")
        chr1$ENSG00000198758 <- chr1$ENSG00000198758.x + chr1$ENSG00000198758.y
        chr1 <- chr1 %>% select(-ENSG00000198758.x, -ENSG00000198758.y)
        cat('n genes in chr 1:', length(chr1) - 1, '\n')
        
        # Filter to your gene set 
        chr1 <- chr1 %>% select(c("Sample_ID", any_of(list)))
        cat('n genes in chr 1 in set:', length(chr1) - 1, '\n')
        # Only proceed if there are genes in the filtered set
        if (ncol(chr1) > 1) {
            chr1[[paste0("chr_1_", var, "_AC", AC, "_total")]] <- rowSums(chr1[,-1])
            geneset <- select(chr1, c("Sample_ID", paste0("chr_1_", var, "_AC", AC, "_total")))
        } else {
            cat("No genes from the gene set found in chr 1.\n")
        }
        rm(chr1)
        rm(first_half)
        rm(second_half)
        gc()
    }
    else {
        # All other chromosomes
        filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC,'_gene_counts.tsv')
        input <- fread(filename, header=TRUE, stringsAsFactors=FALSE)
        print(paste0('Processing chr ', chr))
        cat('n genes in chr:', length(input) - 1, '\n')
        # Filter the columns to keep only the genes present in the gene list
        chr_set <- input %>% select(c("Sample_ID", any_of(list)))
        cat('n genes in set in chr:' , length(chr_set) -1, '\n')
        # Add up variants in genes in set 
        if (ncol(chr_set) > 1) {
            chr_set[[paste0("chr_", chr, "_", var, "_AC", AC, "_total")]] <- rowSums(chr_set[,-1])
            chr_set <- select(chr_set, c("Sample_ID", paste0("chr_", chr, "_", var, "_AC", AC, "_total")))
            
            # Merge to big table and print length (to show increasing in size!!)
            geneset <- merge(geneset, chr_set, by='Sample_ID', all=TRUE)  
            # 'all=TRUE' to retain all rows in case one file is missing data
            cat('n vars in set in chr:', sum(geneset[[paste0("chr_", chr, "_", var, "_AC", AC, "_total")]]), '\n')
        } else {
            cat("No genes from the gene set found in chr", chr, ".\n")
        }
        
        # Clear mem
        rm(input)
        rm(chr_set)
        gc()}
    }

  # Add up across all chromosomes 
  geneset[[paste0(var,"_AC", AC, "_", setname, "_total")]] <- rowSums(geneset[,-1])
  geneset <- select(geneset, c("Sample_ID", paste0(var,"_AC", AC, "_", setname, "_total")))
  print(sum(geneset[[2]]))

  # Save out 
  output <- paste0(setname, '_', var, '_AC', AC, '_counts.csv')
  write.csv(geneset, output)


  upload_command <- paste('dx upload ', output, '--destination ./Variant_counts/Gene_sets/')
  system(upload_command)
}

```

## Regress
### Read in
```{r}
system('dx download ./phenotype_cohorts/processed_cognition/processed_cog_tests_inc_g_Feb24.csv')
phen=read.csv('processed_cog_tests_inc_g_Feb24.csv', header=T)
## Ancestries
system('dx download ./genetic_ancestries/app15175_ids/full_dataset_inferred_ancestries.csv')
ancs_wrongid=read.csv('full_dataset_inferred_ancestries.csv', header=T)
system('dx download ./genetic_ancestries/app15175_ids/PCAs_*')
EUR_PCs_wrongid=read.table('PCAs_EUR_only.tsv', header=T)
AAC_PCs_wrongid=read.table('PCAs_AAC_only.tsv', header=T)
EAS_PCs_wrongid=read.table('PCAs_EAS_only.tsv', header=T)
LAT_PCs_wrongid=read.table('PCAs_LAT_only.tsv', header=T)
SAS_PCs_wrongid=read.table('PCAs_SAS_only.tsv', header=T)
SSA_PCs_wrongid=read.table('PCAs_SSA_only.tsv', header=T)
ADM_PCs_wrongid=read.table('PCAs_ADM_only.tsv', header=T)

system('dx download ./phenotype_cohorts/link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt')
link_ids=read.table('link_AD.app15175_George.app14421_Xav.app17044_JW.app13310.txt', header=T)
ancs=merge(ancs_wrongid, link_ids, by.x='FID', by.y='app15175')
ancs=merge(ancs, EUR_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, AAC_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, EAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, LAT_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SAS_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, SSA_PCs_wrongid, by.x ='FID', by.y='s', all=T)
ancs=merge(ancs, ADM_PCs_wrongid, by.x ='FID', by.y='s', all=T)

# Merge 
phen=merge(phen, ancs, by.x = 'eid', by.y='eid_13310')

## QC
system('dx download ./WES_QC/discordant_sex_ids.csv')
ids_to_rm=read.csv('discordant_sex_ids.csv')
ids_to_rm$remove=T
phen=merge(phen, ids_to_rm, by.x = 'eid', by.y='x', all=T)
phen=subset(phen, is.na(phen$remove)) 


## Covariates
system('dx download ./phenotype_cohorts/covars_participant.csv')
covars=read.csv('covars_participant.csv', header=T)
covars$seq_batch<-covars$p32050
covars$AC_attended<-covars$p54_i0
covars$sex_1male[covars$p31 == 'Male']<-1
covars$sex_1male[covars$p31 == 'Female']<-2
covars$standardised_age <- (covars$p21022 - mean(covars$p21022, na.rm = TRUE)) / sd(covars$p21022, na.rm = TRUE)
covars$standardised_age_squared<-covars$standardised_age*covars$standardised_age
covars$sex_standardised_age<-covars$sex_1male * covars$standardised_age
covars$sex_standardised_age_squared<-covars$sex_1male * covars$standardised_age_squared
phen=merge(phen, covars, by.x='eid', by.y='eid')

## Diagnoses 
system('dx download ./phenotypes_cog_diagnoses_participation_death.csv')
diagnoses=read.csv('phenotypes_cog_diagnoses_participation_death.csv', header=T)
# Remove people with SZ, ID and/or ASD 
diagnoses=subset(diagnoses, diagnoses$SZ==F & diagnoses$ASD==F & diagnoses$intellectual_disability==F) 
diagnoses$no_diagnoses<-T
diagnoses=diagnoses[,c(3,174)]
phen=merge(phen, diagnoses, by='eid') 
```

### Get counts
```{r}
system('dx download ./Variant_counts/Gene_sets/exome_*_AC5orless_*.csv --overwrite')

# REMOVE SINGLE GENE TABLES HERE!!
file_list <- list.files(path = "./", pattern = "*_AC5orless_counts.csv", full.names = TRUE)

# Read, drop first column (V1), and merge each file
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames


all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

### Set covars 
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' # SSA removed here as no one defined as EUR anc has a non-zero SSA probability. 

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

### Prep results table 
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)

sets <- list('SMR_prioritised_negative_expression',
             'SMR_prioritised_positive_expression')

var_types<-list('PTV', 'REVEL75_Miss')
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')
```

### Regress!
```{r}
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

var_types<-list('synonymous')
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

### Save out
```{r}
write.csv(results, 'results_g_regressions_SMR_expressions_FINAL.csv')
system('dx upload results_g_regressions_SMR_expressions_FINAL.csv --destination ./Variant_counts/Gene_sets/reviewer_response/')
```

Download and plot in R!!


# Describe european ancestry probability distribution
```{r}
EU_ancs %>% count(EUR)

EUR1 = subset(EU_ancs, EU_ancs$EUR==1)
EUR1 = subset(EU_ancs, EU_ancs$EUR>0.99)
EUR1 = subset(EU_ancs, EU_ancs$EUR>0.95)
EUR1 = subset(EU_ancs, EU_ancs$EUR>0.90)
```

# Show that not controlling for ancestry probabilities makes no difference
## Get counts
```{r}
system('dx download ./Variant_counts/Gene_sets/pLI*_AC5orless_*.csv --overwrite')

# REMOVE SINGLE GENE TABLES HERE!!
file_list <- list.files(path = "./", pattern = "*_AC5orless_counts.csv", full.names = TRUE)

# Read, drop first column (V1), and merge each file
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames


all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

### Set covars (dropping ancestry probabilities)
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' # SSA removed here as no one defined as EUR anc has a non-zero SSA probability. 

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

### Prep results table 
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)

sets <- list('exome_wide',
              'pLI_0.9')

dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')
```

### Regress!
```{r}
var_types<-list('PTV', 'REVEL75_Miss')
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_no_probabilities'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

var_types<-list('synonymous')
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only_no_probabilities'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

### Save out
```{r}
write.csv(results, 'results_g_regressions_no_probability_covars.csv')
system('dx upload results_g_regressions_no_probability_covars.csv --destination ./Variant_counts/Gene_sets/reviewer_response/')
```



# Show that running in people with an EUR-like probability of 1 makes no difference
## Get counts
```{r}
system('dx download ./Variant_counts/Gene_sets/pLI*_AC5orless_*.csv --overwrite')

# REMOVE SINGLE GENE TABLES HERE!!
file_list <- list.files(path = "./", pattern = "*_AC5orless_counts.csv", full.names = TRUE)

# Read, drop first column (V1), and merge each file
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames


all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
EU_ancs=subset(all_ancs, all_ancs$EUR==1.00) 

### Set covars (dropping ancestry probabilities)
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' # SSA removed here as no one defined as EUR anc has a non-zero SSA probability. 

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

### Prep results table 
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)

sets <- list('exome_wide',
              'pLI_0.9')

dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')
```

### Regress!
```{r}
var_types<-list('PTV', 'REVEL75_Miss')
# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_strict_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

var_types<-list('synonymous')
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_strict_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

### Save out
```{r}
write.csv(results, 'results_g_regressions_strict_EUR.csv')
system('dx upload results_g_regressions_strict_EUR.csv --destination ./Variant_counts/Gene_sets/reviewer_response/')
```


And then plot!! 



# Recounts
## Transposing counts
### AC21 to 50 
run chr 2-16
```{r}
library(data.table)

# Make a list of chromosomes, variant classes, and ACs to process (ensure names match up to table names)
chromosomes <- 1:22
variant_classes <- c("REVEL75_Miss", "PTV", "synonymous")
allele_counts <- '21to50'   

# Loop over each chromosome
for (chr in chromosomes) {
    for (var in variant_classes) {
        for (AC in allele_counts) {
          if (chr == 1) {
            halves <- c("first_half", "second_half")
            for (half in halves) {
                    # Construct the file name for the current half
                    input_filename <- paste0('Variant_counts/untransposed_counts/chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    
                    # Download the file to your project using dx download
                    download_command <- paste('dx download ', shQuote(input_filename))
                    system(download_command)

                    # Read the table for the current half using fread
                    input_filename <- paste0('chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
                     # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
                    counts=subset(counts, !is.na(counts$gene_id)) 
                    
                    # Transpose the table using melt and dcast
                    transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
                    transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")

                    # Write the transposed table out
                    output_filename <- paste0('transposed_chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    fwrite(transposed_counts, output_filename, sep = "\t")

                    # Upload the file to your project using dx upload
                    upload_command <- paste('dx upload', shQuote(output_filename), '--destination Variant_counts/transposed_counts/' )
                    system(upload_command)
                }
            }
          else {
            # Construct the file name 
            input_filename <- paste0('Variant_counts/untransposed_counts/chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')

            # Download the file to your project using dx download
            download_command <- paste('dx download ', shQuote(input_filename))
            system(download_command)
            
            # Read the table for the current chromosome using fread and transpos
            ## Using fread from data table really speeds up the process of reading in 
            input_filename <- paste0('chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
            # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
            counts=subset(counts, !is.na(counts$gene_id)) 
            
            # Transpose the table using melt and dcast:
            ## Melt takes table from wide format into a v long format, where each row represents a single observation (so a single gene and a single sample)
            ### This transposed table is really big so you need a fairly high mem session for this.
            transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
            ## dcast then takes the dataset in this long format and casts it back to a wide format, but this time with sample id as rows and genes as columns
            transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")
            
            # Write the transposed table out 
            output_filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            fwrite(transposed_counts, output_filename, sep = "\t")

            # Upload the file to your project using dx upload
            upload_command <- paste('dx upload', shQuote(output_filename),'--destination Variant_counts/transposed_counts/')
            system(upload_command)
          } 
        }
    }
}
          
```

### AC11 to 20 
```{r}
library(data.table)

# Make a list of chromosomes, variant classes, and ACs to process (ensure names match up to table names)
chromosomes <- 1:22
variant_classes <- c("REVEL75_Miss", "PTV", "synonymous")
allele_counts <- '11to20'   

# Loop over each chromosome
for (chr in chromosomes) {
    for (var in variant_classes) {
        for (AC in allele_counts) {
          if (chr == 1) {
            halves <- c("first_half", "second_half")
            for (half in halves) {
                    # Construct the file name for the current half
                    input_filename <- paste0('Variant_counts/untransposed_counts/chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    
                    # Download the file to your project using dx download
                    download_command <- paste('dx download ', shQuote(input_filename))
                    system(download_command)

                    # Read the table for the current half using fread
                    input_filename <- paste0('chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
                     # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
                    counts=subset(counts, !is.na(counts$gene_id)) 
                    
                    # Transpose the table using melt and dcast
                    transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
                    transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")

                    # Write the transposed table out
                    output_filename <- paste0('transposed_chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    fwrite(transposed_counts, output_filename, sep = "\t")

                    # Upload the file to your project using dx upload
                    upload_command <- paste('dx upload', shQuote(output_filename), '--destination Variant_counts/transposed_counts/' )
                    system(upload_command)
                }
            }
          else {
            # Construct the file name 
            input_filename <- paste0('Variant_counts/untransposed_counts/chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')

            # Download the file to your project using dx download
            download_command <- paste('dx download ', shQuote(input_filename))
            system(download_command)
            
            # Read the table for the current chromosome using fread and transpos
            ## Using fread from data table really speeds up the process of reading in 
            input_filename <- paste0('chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
            # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
            counts=subset(counts, !is.na(counts$gene_id)) 
            
            # Transpose the table using melt and dcast:
            ## Melt takes table from wide format into a v long format, where each row represents a single observation (so a single gene and a single sample)
            ### This transposed table is really big so you need a fairly high mem session for this.
            transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
            ## dcast then takes the dataset in this long format and casts it back to a wide format, but this time with sample id as rows and genes as columns
            transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")
            
            # Write the transposed table out 
            output_filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            fwrite(transposed_counts, output_filename, sep = "\t")

            # Upload the file to your project using dx upload
            upload_command <- paste('dx upload', shQuote(output_filename),'--destination Variant_counts/transposed_counts/')
            system(upload_command)
          } 
        }
    }
}
          
```

### AC6 to 10 
```{r}
library(data.table)

# Make a list of chromosomes, variant classes, and ACs to process (ensure names match up to table names)
chromosomes <- 1:22
variant_classes <- c("REVEL75_Miss", "PTV", "synonymous")
allele_counts <- '6to10'   

# Loop over each chromosome
for (chr in chromosomes) {
    for (var in variant_classes) {
        for (AC in allele_counts) {
          if (chr == 1) {
            halves <- c("first_half", "second_half")
            for (half in halves) {
                    # Construct the file name for the current half
                    input_filename <- paste0('Variant_counts/untransposed_counts/chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    
                    # Download the file to your project using dx download
                    download_command <- paste('dx download ', shQuote(input_filename))
                    system(download_command)

                    # Read the table for the current half using fread
                    input_filename <- paste0('chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
                     # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
                    counts=subset(counts, !is.na(counts$gene_id)) 
                    
                    # Transpose the table using melt and dcast
                    transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
                    transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")

                    # Write the transposed table out
                    output_filename <- paste0('transposed_chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    fwrite(transposed_counts, output_filename, sep = "\t")

                    # Upload the file to your project using dx upload
                    upload_command <- paste('dx upload', shQuote(output_filename), '--destination Variant_counts/transposed_counts/' )
                    system(upload_command)
                }
            }
          else {
            # Construct the file name 
            input_filename <- paste0('Variant_counts/untransposed_counts/chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')

            # Download the file to your project using dx download
            download_command <- paste('dx download ', shQuote(input_filename))
            system(download_command)
            
            # Read the table for the current chromosome using fread and transpos
            ## Using fread from data table really speeds up the process of reading in 
            input_filename <- paste0('chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
            # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
            counts=subset(counts, !is.na(counts$gene_id)) 
            
            # Transpose the table using melt and dcast:
            ## Melt takes table from wide format into a v long format, where each row represents a single observation (so a single gene and a single sample)
            ### This transposed table is really big so you need a fairly high mem session for this.
            transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
            ## dcast then takes the dataset in this long format and casts it back to a wide format, but this time with sample id as rows and genes as columns
            transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")
            
            # Write the transposed table out 
            output_filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            fwrite(transposed_counts, output_filename, sep = "\t")

            # Upload the file to your project using dx upload
            upload_command <- paste('dx upload', shQuote(output_filename),'--destination Variant_counts/transposed_counts/')
            system(upload_command)
          } 
        }
    }
}
          
```

### Less deleterious missense (AC 5 or less)
```{r}
library(data.table)

# Make a list of chromosomes, variant classes, and ACs to process (ensure names match up to table names)
chromosomes <- 1:22
variant_classes <- c("benign_REVEL50_Miss", "mid_REVEL50to75_Miss")
allele_counts <- '5orless'   

# Loop over each chromosome
for (chr in chromosomes) {
    for (var in variant_classes) {
        for (AC in allele_counts) {
          if (chr == 1) {
            halves <- c("first_half", "second_half")
            for (half in halves) {
                    # Construct the file name for the current half
                    input_filename <- paste0('Variant_counts/untransposed_counts/chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    
                    # Download the file to your project using dx download
                    download_command <- paste('dx download ', shQuote(input_filename))
                    system(download_command)

                    # Read the table for the current half using fread
                    input_filename <- paste0('chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
                     # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
                    counts=subset(counts, !is.na(counts$gene_id)) 
                    
                    # Transpose the table using melt and dcast
                    transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
                    transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")

                    # Write the transposed table out
                    output_filename <- paste0('transposed_chr_', chr, '_', half, '_', var, '_AC', AC, '_gene_counts.tsv')
                    fwrite(transposed_counts, output_filename, sep = "\t")

                    # Upload the file to your project using dx upload
                    upload_command <- paste('dx upload', shQuote(output_filename), '--destination Variant_counts/transposed_counts/' )
                    system(upload_command)
                }
            }
          else {
            # Construct the file name 
            input_filename <- paste0('Variant_counts/untransposed_counts/chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')

            # Download the file to your project using dx download
            download_command <- paste('dx download ', shQuote(input_filename))
            system(download_command)
            
            # Read the table for the current chromosome using fread and transpos
            ## Using fread from data table really speeds up the process of reading in 
            input_filename <- paste0('chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            counts <- fread(input_filename, header=TRUE, stringsAsFactors=FALSE)
            # Remove the gene_id that is NA - if there are any variants not assigned to a gene then Hail writes them to here in the counting process. 
            counts=subset(counts, !is.na(counts$gene_id)) 
            
            # Transpose the table using melt and dcast:
            ## Melt takes table from wide format into a v long format, where each row represents a single observation (so a single gene and a single sample)
            ### This transposed table is really big so you need a fairly high mem session for this.
            transposed_counts <- melt(counts, id.vars = "gene_id", variable.name = "Sample_ID", value.name = "Burden")
            ## dcast then takes the dataset in this long format and casts it back to a wide format, but this time with sample id as rows and genes as columns
            transposed_counts <- dcast(transposed_counts, Sample_ID ~ gene_id, value.var = "Burden")
            
            # Write the transposed table out 
            output_filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv')
            fwrite(transposed_counts, output_filename, sep = "\t")

            # Upload the file to your project using dx upload
            upload_command <- paste('dx upload', shQuote(output_filename),'--destination Variant_counts/transposed_counts/')
            system(upload_command)
          } 
        }
    }
}
          
```

## Gene set counts 
### AC21to50
```{r}
library(data.table)
library(dplyr)
AC='21to50' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('PTV', 'REVEL75_Miss', 'synonymous')

for (var in variant_classes) {
  input_filename <- sprintf('./Variant_counts/transposed_counts/AC%s/AC%s_%s/transposed*.tsv', AC, AC, var)
  download_command <- sprintf('dx download %s', shQuote(input_filename))
  system(download_command)
}


# Gene set annotations
system('dx download ./gene_sets_for_RAP_Feb25.csv')
sets=read.csv('gene_sets_for_RAP_Feb25.csv', header=T)


# Gene set definitions
gene_sets <- list(
  pLI_0.9 = subset(sets, sets$pLI_constrained == TRUE)$gene_id,
  non_constrained_pLI_lessthan0.9 = subset(sets, sets$pLI_constrained == FALSE)$gene_id,
  brain_expressed = subset(sets, sets$Brain_expressed == TRUE)$gene_id,
  non_brain_expressed = subset(sets, sets$Brain_expressed == F)$gene_id,
  SCHEMA_FDR_US = subset(sets, sets$SCHEMA_FDR5_gene == TRUE)$gene_id,
  SZ_GWAS_all = subset(sets, sets$SZ_GWAS_gene == TRUE)$gene_id,
  SZ_GWAS_closest = subset(sets, sets$SZ_GWAS_closest_indexed_SNP == TRUE)$gene_id,
  SZ_GWAS_credible_causal_US = subset(sets, sets$SZ_GWAS_credible_causal == TRUE)$gene_id,
  NDD_CNVs_US = subset(sets, sets$NDD_CNV_gene == TRUE)$gene_id,
  SZ_CNVs_US = subset(sets, sets$SZ_CNV_gene == TRUE)$gene_id)

for (setname in names(gene_sets)) {
  cat("\n### Processing gene set:", setname, "###\n")
  gene_list <- gene_sets[[setname]]

  for (var in variant_classes) {
    cat("\nRunning variant class:", var, "\n")
    geneset <- NULL  # initialise output

    for (chr in chromosomes) {
      cat("Processing chr", chr, "\n")

      # Read and merge chr 1 halves
      if (chr == 1) {
        first_half <- fread(paste0('transposed_chr_1_first_half_', var, '_AC', AC, '_gene_counts.tsv'))
        second_half <- fread(paste0('transposed_chr_1_second_half_', var, '_AC', AC, '_gene_counts.tsv'))

        chr_data <- merge(first_half, second_half, by = "Sample_ID")
        # Sum over duplicated gene (specific to ENSG00000198758)
        chr_data$ENSG00000198758 <- chr_data$ENSG00000198758.x + chr_data$ENSG00000198758.y
        chr_data <- chr_data %>% select(-ENSG00000198758.x, -ENSG00000198758.y)

        # Clean up
        rm(first_half, second_half)
      } else {
        chr_data <- fread(paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv'))
      }

      # Filter to gene list if set defined (skip for exome_wide)
      if (!is.null(gene_list)) {
        chr_data <- chr_data %>% select(c("Sample_ID", any_of(gene_list)))
      }

      # Skip empty sets
      if (ncol(chr_data) <= 1) {
        cat("No genes in this set for chr", chr, "\n")
        rm(chr_data)
        gc()
        next
      }

      # Add chromosome total
      chr_total_col <- paste0("chr_", chr, "_", var, "_AC", AC, "_total")
      chr_data[[chr_total_col]] <- rowSums(chr_data[,-1], na.rm = TRUE)
      chr_data <- chr_data %>% select("Sample_ID", chr_total_col)

      # Merge
      if (is.null(geneset)) {
        geneset <- chr_data
      } else {
        geneset <- merge(geneset, chr_data, by = "Sample_ID", all = TRUE)
      }

      rm(chr_data)
      gc()
    }

    # Final total per gene set
    total_col <- paste0(var, "_AC", AC, "_", setname, "_total")
    geneset[[total_col]] <- rowSums(geneset[,-1], na.rm = TRUE)
    geneset <- select(geneset, c("Sample_ID", total_col))
    cat("Total variants across all chromosomes:", sum(geneset[[2]], na.rm = TRUE), "\n")

    # Save + upload
    output <- paste0(setname, "_", var, "_AC", AC, "_counts.csv")
    write.csv(geneset, output, row.names = FALSE)

    upload_command <- paste("dx upload", output, "--destination ./Variant_counts/Gene_sets/")
    system(upload_command)

    cat("Completed:", output, "\n")
  }
}

```

### AC11to20
```{r}
library(data.table)
library(dplyr)
AC='11to20' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('PTV', 'REVEL75_Miss', 'synonymous')

for (var in variant_classes) {
  input_filename <- sprintf('./Variant_counts/transposed_counts/AC%s/AC%s_%s/transposed*.tsv', AC, AC, var)
  download_command <- sprintf('dx download %s', shQuote(input_filename))
  system(download_command)
}


# Gene set annotations
system('dx download ./gene_sets_for_RAP_Feb25.csv')
sets=read.csv('gene_sets_for_RAP_Feb25.csv', header=T)


# Gene set definitions
gene_sets <- list(
  pLI_0.9 = subset(sets, sets$pLI_constrained == TRUE)$gene_id,
  non_constrained_pLI_lessthan0.9 = subset(sets, sets$pLI_constrained == FALSE)$gene_id,
  brain_expressed = subset(sets, sets$Brain_expressed == TRUE)$gene_id,
  non_brain_expressed = subset(sets, sets$Brain_expressed == F)$gene_id,
  SCHEMA_FDR_US = subset(sets, sets$SCHEMA_FDR5_gene == TRUE)$gene_id,
  SZ_GWAS_all = subset(sets, sets$SZ_GWAS_gene == TRUE)$gene_id,
  SZ_GWAS_closest = subset(sets, sets$SZ_GWAS_closest_indexed_SNP == TRUE)$gene_id,
  SZ_GWAS_credible_causal_US = subset(sets, sets$SZ_GWAS_credible_causal == TRUE)$gene_id,
  NDD_CNVs_US = subset(sets, sets$NDD_CNV_gene == TRUE)$gene_id,
  SZ_CNVs_US = subset(sets, sets$SZ_CNV_gene == TRUE)$gene_id)

for (setname in names(gene_sets)) {
  cat("\n### Processing gene set:", setname, "###\n")
  gene_list <- gene_sets[[setname]]

  for (var in variant_classes) {
    cat("\nRunning variant class:", var, "\n")
    geneset <- NULL  # initialise output

    for (chr in chromosomes) {
      cat("Processing chr", chr, "\n")

      # Read and merge chr 1 halves
      if (chr == 1) {
        first_half <- fread(paste0('transposed_chr_1_first_half_', var, '_AC', AC, '_gene_counts.tsv'))
        second_half <- fread(paste0('transposed_chr_1_second_half_', var, '_AC', AC, '_gene_counts.tsv'))

        chr_data <- merge(first_half, second_half, by = "Sample_ID")
        # Sum over duplicated gene (specific to ENSG00000198758)
        chr_data$ENSG00000198758 <- chr_data$ENSG00000198758.x + chr_data$ENSG00000198758.y
        chr_data <- chr_data %>% select(-ENSG00000198758.x, -ENSG00000198758.y)

        # Clean up
        rm(first_half, second_half)
      } else {
        chr_data <- fread(paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv'))
      }

      # Filter to gene list if set defined (skip for exome_wide)
      if (!is.null(gene_list)) {
        chr_data <- chr_data %>% select(c("Sample_ID", any_of(gene_list)))
      }

      # Skip empty sets
      if (ncol(chr_data) <= 1) {
        cat("No genes in this set for chr", chr, "\n")
        rm(chr_data)
        gc()
        next
      }

      # Add chromosome total
      chr_total_col <- paste0("chr_", chr, "_", var, "_AC", AC, "_total")
      chr_data[[chr_total_col]] <- rowSums(chr_data[,-1], na.rm = TRUE)
      chr_data <- chr_data %>% select("Sample_ID", chr_total_col)

      # Merge
      if (is.null(geneset)) {
        geneset <- chr_data
      } else {
        geneset <- merge(geneset, chr_data, by = "Sample_ID", all = TRUE)
      }

      rm(chr_data)
      gc()
    }

    # Final total per gene set
    total_col <- paste0(var, "_AC", AC, "_", setname, "_total")
    geneset[[total_col]] <- rowSums(geneset[,-1], na.rm = TRUE)
    geneset <- select(geneset, c("Sample_ID", total_col))
    cat("Total variants across all chromosomes:", sum(geneset[[2]], na.rm = TRUE), "\n")

    # Save + upload
    output <- paste0(setname, "_", var, "_AC", AC, "_counts.csv")
    write.csv(geneset, output, row.names = FALSE)

    upload_command <- paste("dx upload", output, "--destination ./Variant_counts/Gene_sets/")
    system(upload_command)

    cat("Completed:", output, "\n")
  }
}

```

### AC6to10
```{r}
library(data.table)
library(dplyr)
AC='6to10' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('PTV', 'REVEL75_Miss', 'synonymous')

for (var in variant_classes) {
  input_filename <- sprintf('./Variant_counts/transposed_counts/AC%s/AC%s_%s/transposed*.tsv', AC, AC, var)
  download_command <- sprintf('dx download %s', shQuote(input_filename))
  system(download_command)
}


# Gene set annotations
system('dx download ./gene_sets_for_RAP_Feb25.csv')
sets=read.csv('gene_sets_for_RAP_Feb25.csv', header=T)


# Gene set definitions
gene_sets <- list(
  pLI_0.9 = subset(sets, sets$pLI_constrained == TRUE)$gene_id,
  non_constrained_pLI_lessthan0.9 = subset(sets, sets$pLI_constrained == FALSE)$gene_id,
  brain_expressed = subset(sets, sets$Brain_expressed == TRUE)$gene_id,
  non_brain_expressed = subset(sets, sets$Brain_expressed == F)$gene_id,
  SCHEMA_FDR_US = subset(sets, sets$SCHEMA_FDR5_gene == TRUE)$gene_id,
  SZ_GWAS_all = subset(sets, sets$SZ_GWAS_gene == TRUE)$gene_id,
  SZ_GWAS_closest = subset(sets, sets$SZ_GWAS_closest_indexed_SNP == TRUE)$gene_id,
  SZ_GWAS_credible_causal_US = subset(sets, sets$SZ_GWAS_credible_causal == TRUE)$gene_id,
  NDD_CNVs_US = subset(sets, sets$NDD_CNV_gene == TRUE)$gene_id,
  SZ_CNVs_US = subset(sets, sets$SZ_CNV_gene == TRUE)$gene_id)

for (setname in names(gene_sets)) {
  cat("\n### Processing gene set:", setname, "###\n")
  gene_list <- gene_sets[[setname]]

  for (var in variant_classes) {
    cat("\nRunning variant class:", var, "\n")
    geneset <- NULL  # initialise output

    for (chr in chromosomes) {
      cat("Processing chr", chr, "\n")

      # Read and merge chr 1 halves
      if (chr == 1) {
        first_half <- fread(paste0('transposed_chr_1_first_half_', var, '_AC', AC, '_gene_counts.tsv'))
        second_half <- fread(paste0('transposed_chr_1_second_half_', var, '_AC', AC, '_gene_counts.tsv'))

        chr_data <- merge(first_half, second_half, by = "Sample_ID")
        # Sum over duplicated gene (specific to ENSG00000198758)
        chr_data$ENSG00000198758 <- chr_data$ENSG00000198758.x + chr_data$ENSG00000198758.y
        chr_data <- chr_data %>% select(-ENSG00000198758.x, -ENSG00000198758.y)

        # Clean up
        rm(first_half, second_half)
      } else {
        chr_data <- fread(paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv'))
      }

      # Filter to gene list if set defined (skip for exome_wide)
      if (!is.null(gene_list)) {
        chr_data <- chr_data %>% select(c("Sample_ID", any_of(gene_list)))
      }

      # Skip empty sets
      if (ncol(chr_data) <= 1) {
        cat("No genes in this set for chr", chr, "\n")
        rm(chr_data)
        gc()
        next
      }

      # Add chromosome total
      chr_total_col <- paste0("chr_", chr, "_", var, "_AC", AC, "_total")
      chr_data[[chr_total_col]] <- rowSums(chr_data[,-1], na.rm = TRUE)
      chr_data <- chr_data %>% select("Sample_ID", chr_total_col)

      # Merge
      if (is.null(geneset)) {
        geneset <- chr_data
      } else {
        geneset <- merge(geneset, chr_data, by = "Sample_ID", all = TRUE)
      }

      rm(chr_data)
      gc()
    }

    # Final total per gene set
    total_col <- paste0(var, "_AC", AC, "_", setname, "_total")
    geneset[[total_col]] <- rowSums(geneset[,-1], na.rm = TRUE)
    geneset <- select(geneset, c("Sample_ID", total_col))
    cat("Total variants across all chromosomes:", sum(geneset[[2]], na.rm = TRUE), "\n")

    # Save + upload
    output <- paste0(setname, "_", var, "_AC", AC, "_counts.csv")
    write.csv(geneset, output, row.names = FALSE)

    upload_command <- paste("dx upload", output, "--destination ./Variant_counts/Gene_sets/")
    system(upload_command)

    cat("Completed:", output, "\n")
  }
}

```

### Less deleterious missense 
```{r}
library(data.table)
library(dplyr)
AC='5orless' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('mid_REVEL50to75_Miss', 'benign_REVEL50_Miss')

for (var in variant_classes) {
  input_filename <- sprintf('Variant_counts/transposed_counts/AC%s/AC%s_%s/transposed*.tsv', AC, AC, var)
  download_command <- sprintf('dx download %s', shQuote(input_filename))
  system(download_command)
}


# Gene set annotations
system('dx download ./gene_sets_for_RAP_Feb25.csv')
sets=read.csv('gene_sets_for_RAP_Feb25.csv', header=T)


# Gene set definitions
gene_sets <- list(
  pLI_0.9 = subset(sets, sets$pLI_constrained == TRUE)$gene_id,
  non_constrained_pLI_lessthan0.9 = subset(sets, sets$pLI_constrained == FALSE)$gene_id,
  brain_expressed = subset(sets, sets$Brain_expressed == TRUE)$gene_id,
  non_brain_expressed = subset(sets, sets$Brain_expressed == F)$gene_id,
  SCHEMA_FDR_US = subset(sets, sets$SCHEMA_FDR5_gene == TRUE)$gene_id,
  SZ_GWAS_all = subset(sets, sets$SZ_GWAS_gene == TRUE)$gene_id,
  SZ_GWAS_closest = subset(sets, sets$SZ_GWAS_closest_indexed_SNP == TRUE)$gene_id,
  SZ_GWAS_credible_causal_US = subset(sets, sets$SZ_GWAS_credible_causal == TRUE)$gene_id,
  NDD_CNVs_US = subset(sets, sets$NDD_CNV_gene == TRUE)$gene_id,
  SZ_CNVs_US = subset(sets, sets$SZ_CNV_gene == TRUE)$gene_id)

for (setname in names(gene_sets)) {
  cat("\n### Processing gene set:", setname, "###\n")
  gene_list <- gene_sets[[setname]]

  for (var in variant_classes) {
    cat("\nRunning variant class:", var, "\n")
    geneset <- NULL  # initialise output

    for (chr in chromosomes) {
      cat("Processing chr", chr, "\n")

      # Read and merge chr 1 halves
      if (chr == 1) {
        first_half <- fread(paste0('transposed_chr_1_first_half_', var, '_AC', AC, '_gene_counts.tsv'))
        second_half <- fread(paste0('transposed_chr_1_second_half_', var, '_AC', AC, '_gene_counts.tsv'))

        chr_data <- merge(first_half, second_half, by = "Sample_ID")
        # Sum over duplicated gene (specific to ENSG00000198758)
        chr_data$ENSG00000198758 <- chr_data$ENSG00000198758.x + chr_data$ENSG00000198758.y
        chr_data <- chr_data %>% select(-ENSG00000198758.x, -ENSG00000198758.y)

        # Clean up
        rm(first_half, second_half)
      } else {
        chr_data <- fread(paste0('transposed_chr_', chr, '_', var, '_AC', AC, '_gene_counts.tsv'))
      }

      # Filter to gene list if set defined (skip for exome_wide)
      if (!is.null(gene_list)) {
        chr_data <- chr_data %>% select(c("Sample_ID", any_of(gene_list)))
      }

      # Skip empty sets
      if (ncol(chr_data) <= 1) {
        cat("No genes in this set for chr", chr, "\n")
        rm(chr_data)
        gc()
        next
      }

      # Add chromosome total
      chr_total_col <- paste0("chr_", chr, "_", var, "_AC", AC, "_total")
      chr_data[[chr_total_col]] <- rowSums(chr_data[,-1], na.rm = TRUE)
      chr_data <- chr_data %>% select("Sample_ID", chr_total_col)

      # Merge
      if (is.null(geneset)) {
        geneset <- chr_data
      } else {
        geneset <- merge(geneset, chr_data, by = "Sample_ID", all = TRUE)
      }

      rm(chr_data)
      gc()
    }

    # Final total per gene set
    total_col <- paste0(var, "_AC", AC, "_", setname, "_total")
    geneset[[total_col]] <- rowSums(geneset[,-1], na.rm = TRUE)
    geneset <- select(geneset, c("Sample_ID", total_col))
    cat("Total variants across all chromosomes:", sum(geneset[[2]], na.rm = TRUE), "\n")

    # Save + upload
    output <- paste0(setname, "_", var, "_AC", AC, "_counts.csv")
    write.csv(geneset, output, row.names = FALSE)

    upload_command <- paste("dx upload", output, "--destination ./Variant_counts/Gene_sets/")
    system(upload_command)

    cat("Completed:", output, "\n")
  }
}

```

### Exome wide counts for ACs>5
#### 21to50
```{r}
library(data.table)
library(dplyr)
AC='21to50' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('PTV', 'REVEL75_Miss', 'synonymous')

for (var in variant_classes) {
  input_filename <- sprintf('./Variant_counts/transposed_counts/AC%s/AC%s_%s/transposed*.tsv', AC, AC, var)
  download_command <- sprintf('dx download %s', shQuote(input_filename))
  system(download_command)
}


setname='exome_wide'
for (var in variant_classes){
  print(paste0('Running variant class ', var))
  for (chr in chromosomes) {
    # Process chr1 differently (as it's in 2 halves)
      if (chr==1) {
        # Read in 
        first_half <- paste0('transposed_chr_1_first_half_', var, '_AC', AC,'_gene_counts.tsv')
        first_half <- fread(first_half, header=TRUE, stringsAsFactors=FALSE)
        second_half <- paste0('transposed_chr_1_second_half_', var, '_AC', AC,'_gene_counts.tsv')
        second_half <- fread(second_half, header=TRUE, stringsAsFactors=FALSE)

        # Merge to one table and add up the gene split across the two halves
        chr1 <- merge(first_half, second_half, by = "Sample_ID")
        chr1$ENSG00000198758 <- chr1$ENSG00000198758.x + chr1$ENSG00000198758.y
        chr1 <- chr1 %>% select(-ENSG00000198758.x, -ENSG00000198758.y)
        cat('n genes in chr 1:', length(chr1) - 1, '\n')
        

        chr1[[paste0("chr_1_", var,"_AC", AC, "_total")]] <- rowSums(chr1[,-1])
        geneset<-select(chr1, c("Sample_ID", paste0("chr_1_", var,"_AC", AC, "_total")))
        rm(chr1)
        rm(first_half)
        rm(second_half)
        gc()
      }
     else {
        # All other chromosomes
        filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC,'_gene_counts.tsv')
        input <- fread(filename, header=TRUE, stringsAsFactors=FALSE)
        print(paste0('Processing chr ', chr))
        cat('n genes in chr:', length(input) - 1, '\n')
        chr_set<-input
        # Add up variants in genes in set 
        chr_set[[paste0("chr_", chr, "_", var,"_AC", AC, "_total")]] <- rowSums(chr_set[,-1])
        chr_set<-select(chr_set, c("Sample_ID", paste0("chr_",chr, "_", var,"_AC", AC, "_total")))
        
        # Merge to big table and print length (to show increasing in size!!)
        geneset<-merge(geneset, chr_set, by='Sample_ID')
        cat('n vars in set in chr:', sum(geneset[[paste0("chr_", chr, "_", var,"_AC", AC, "_total")]]))
        
        # Clear mem
        rm(input)
        rm(chr_set)
        gc()}}

  # Add up across all chromosomes 
  geneset[[paste0(var,"_AC", AC, "_", setname, "_total")]] <- rowSums(geneset[,-1])
  geneset <- select(geneset, c("Sample_ID", paste0(var,"_AC", AC, "_", setname, "_total")))
  print(sum(geneset[[2]]))

  # Save out 
  output <- paste0(setname, '_', var, '_AC', AC, '_counts.csv')
  write.csv(geneset, output)


  upload_command <- paste('dx upload ', output, '--destination ./Variant_counts/Gene_sets/')
  system(upload_command)
}
```

#### 11to20
```{r}
library(data.table)
library(dplyr)
AC='11to20' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('PTV', 'REVEL75_Miss', 'synonymous')

for (var in variant_classes) {
  input_filename <- sprintf('./Variant_counts/transposed_counts/AC%s/AC%s_%s/transposed*.tsv', AC, AC, var)
  download_command <- sprintf('dx download %s', shQuote(input_filename))
  system(download_command)
}


setname='exome_wide'
for (var in variant_classes){
  print(paste0('Running variant class ', var))
  for (chr in chromosomes) {
    # Process chr1 differently (as it's in 2 halves)
      if (chr==1) {
        # Read in 
        first_half <- paste0('transposed_chr_1_first_half_', var, '_AC', AC,'_gene_counts.tsv')
        first_half <- fread(first_half, header=TRUE, stringsAsFactors=FALSE)
        second_half <- paste0('transposed_chr_1_second_half_', var, '_AC', AC,'_gene_counts.tsv')
        second_half <- fread(second_half, header=TRUE, stringsAsFactors=FALSE)

        # Merge to one table and add up the gene split across the two halves
        chr1 <- merge(first_half, second_half, by = "Sample_ID")
        chr1$ENSG00000198758 <- chr1$ENSG00000198758.x + chr1$ENSG00000198758.y
        chr1 <- chr1 %>% select(-ENSG00000198758.x, -ENSG00000198758.y)
        cat('n genes in chr 1:', length(chr1) - 1, '\n')
        

        chr1[[paste0("chr_1_", var,"_AC", AC, "_total")]] <- rowSums(chr1[,-1])
        geneset<-select(chr1, c("Sample_ID", paste0("chr_1_", var,"_AC", AC, "_total")))
        rm(chr1)
        rm(first_half)
        rm(second_half)
        gc()
      }
     else {
        # All other chromosomes
        filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC,'_gene_counts.tsv')
        input <- fread(filename, header=TRUE, stringsAsFactors=FALSE)
        print(paste0('Processing chr ', chr))
        cat('n genes in chr:', length(input) - 1, '\n')
        chr_set<-input
        # Add up variants in genes in set 
        chr_set[[paste0("chr_", chr, "_", var,"_AC", AC, "_total")]] <- rowSums(chr_set[,-1])
        chr_set<-select(chr_set, c("Sample_ID", paste0("chr_",chr, "_", var,"_AC", AC, "_total")))
        
        # Merge to big table and print length (to show increasing in size!!)
        geneset<-merge(geneset, chr_set, by='Sample_ID')
        cat('n vars in set in chr:', sum(geneset[[paste0("chr_", chr, "_", var,"_AC", AC, "_total")]]))
        
        # Clear mem
        rm(input)
        rm(chr_set)
        gc()}}

  # Add up across all chromosomes 
  geneset[[paste0(var,"_AC", AC, "_", setname, "_total")]] <- rowSums(geneset[,-1])
  geneset <- select(geneset, c("Sample_ID", paste0(var,"_AC", AC, "_", setname, "_total")))
  print(sum(geneset[[2]]))

  # Save out 
  output <- paste0(setname, '_', var, '_AC', AC, '_counts.csv')
  write.csv(geneset, output)


  upload_command <- paste('dx upload ', output, '--destination ./Variant_counts/Gene_sets/')
  system(upload_command)
}
```


#### 6to10
```{r}
library(data.table)
library(dplyr)
AC='6to10' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('PTV', 'REVEL75_Miss', 'synonymous')

for (var in variant_classes) {
  input_filename <- sprintf('./Variant_counts/transposed_counts/AC%s/AC%s_%s/transposed*.tsv', AC, AC, var)
  download_command <- sprintf('dx download %s', shQuote(input_filename))
  system(download_command)
}


setname='exome_wide'
for (var in variant_classes){
  print(paste0('Running variant class ', var))
  for (chr in chromosomes) {
    # Process chr1 differently (as it's in 2 halves)
      if (chr==1) {
        # Read in 
        first_half <- paste0('transposed_chr_1_first_half_', var, '_AC', AC,'_gene_counts.tsv')
        first_half <- fread(first_half, header=TRUE, stringsAsFactors=FALSE)
        second_half <- paste0('transposed_chr_1_second_half_', var, '_AC', AC,'_gene_counts.tsv')
        second_half <- fread(second_half, header=TRUE, stringsAsFactors=FALSE)

        # Merge to one table and add up the gene split across the two halves
        chr1 <- merge(first_half, second_half, by = "Sample_ID")
        chr1$ENSG00000198758 <- chr1$ENSG00000198758.x + chr1$ENSG00000198758.y
        chr1 <- chr1 %>% select(-ENSG00000198758.x, -ENSG00000198758.y)
        cat('n genes in chr 1:', length(chr1) - 1, '\n')
        

        chr1[[paste0("chr_1_", var,"_AC", AC, "_total")]] <- rowSums(chr1[,-1])
        geneset<-select(chr1, c("Sample_ID", paste0("chr_1_", var,"_AC", AC, "_total")))
        rm(chr1)
        rm(first_half)
        rm(second_half)
        gc()
      }
     else {
        # All other chromosomes
        filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC,'_gene_counts.tsv')
        input <- fread(filename, header=TRUE, stringsAsFactors=FALSE)
        print(paste0('Processing chr ', chr))
        cat('n genes in chr:', length(input) - 1, '\n')
        chr_set<-input
        # Add up variants in genes in set 
        chr_set[[paste0("chr_", chr, "_", var,"_AC", AC, "_total")]] <- rowSums(chr_set[,-1])
        chr_set<-select(chr_set, c("Sample_ID", paste0("chr_",chr, "_", var,"_AC", AC, "_total")))
        
        # Merge to big table and print length (to show increasing in size!!)
        geneset<-merge(geneset, chr_set, by='Sample_ID')
        cat('n vars in set in chr:', sum(geneset[[paste0("chr_", chr, "_", var,"_AC", AC, "_total")]]))
        
        # Clear mem
        rm(input)
        rm(chr_set)
        gc()}}

  # Add up across all chromosomes 
  geneset[[paste0(var,"_AC", AC, "_", setname, "_total")]] <- rowSums(geneset[,-1])
  geneset <- select(geneset, c("Sample_ID", paste0(var,"_AC", AC, "_", setname, "_total")))
  print(sum(geneset[[2]]))

  # Save out 
  output <- paste0(setname, '_', var, '_AC', AC, '_counts.csv')
  write.csv(geneset, output)


  upload_command <- paste('dx upload ', output, '--destination ./Variant_counts/Gene_sets/')
  system(upload_command)
}
```

## Regressions
### Set up
```{r}
library(data.table)
library(dplyr)
library(purrr)
install.packages('broom')
library(broom)
library(ggplot2)

# READ IN
# Phenotypes
system('dx download ./phenotype_cohorts/processed_cognition/processed_cog_tests_inc_g_Feb24.csv')
phen=read.csv('processed_cog_tests_inc_g_Feb24.csv', header=T)
# Ancestries
system('dx download ./genetic_ancestries/ancestries_PCs.csv')
ancs = read.csv('ancestries_PCs.csv')
phen=merge(phen, ancs, by.x = 'eid', by.y='eid_13310')
# QC
system('dx download ./WES_QC/discordant_sex_ids.csv')
ids_to_rm=read.csv('discordant_sex_ids.csv')
ids_to_rm$remove=T
phen=merge(phen, ids_to_rm, by.x = 'eid', by.y='x', all=T)
phen=subset(phen, is.na(phen$remove)) 
# Covars
system('dx download ./phenotype_cohorts/covars_participant.csv')
covars=read.csv('covars_participant.csv', header=T)
covars$seq_batch<-covars$p32050
covars$AC_attended<-covars$p54_i0
covars$sex_1male[covars$p31 == 'Male']<-1
covars$sex_1male[covars$p31 == 'Female']<-2
covars$standardised_age <- (covars$p21022 - mean(covars$p21022, na.rm = TRUE)) / sd(covars$p21022, na.rm = TRUE)
covars$standardised_age_squared<-covars$standardised_age*covars$standardised_age
covars$sex_standardised_age<-covars$sex_1male * covars$standardised_age
covars$sex_standardised_age_squared<-covars$sex_1male * covars$standardised_age_squared
phen=merge(phen, covars, by.x='eid', by.y='eid')
# Diagnoses
system('dx download ./phenotypes_cog_diagnoses_participation_death.csv')
diagnoses=read.csv('phenotypes_cog_diagnoses_participation_death.csv', header=T)
# Remove people with SZ, ID and/or ASD 
diagnoses=subset(diagnoses, diagnoses$SZ==F & diagnoses$ASD==F & diagnoses$intellectual_disability==F) 
diagnoses$no_diagnoses<-T
diagnoses=diagnoses[,c(3,174)]
phen=merge(phen, diagnoses, by='eid') 
# EUR ancs
all_ancs=phen
all_ancs_bu = all_ancs

```

### AC5 or less - mid and benign missense 
```{r}
# Gene set counts
AC='5orless'
system('dx download ./Variant_counts/Gene_sets/*_AC5orless_*.csv --overwrite')
file_list <- list.files(path = "./", pattern = "*_AC5orless_counts.csv", full.names = TRUE)
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames
all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

# SET UP
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' # SSA removed here as no one defined as EUR anc has a non-zero SSA probability. 
covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE)


# REGRESS
# Non-synonymous vars 
sets <- list('pLI_0.9','non_constrained_pLI_lessthan0.9',
             'brain_expressed','non_brain_expressed',
             'SZ_GWAS_all', 'SZ_GWAS_closest', 'SZ_GWAS_credible_causal_US',
             'NDD_CNVs_US', 'SZ_CNVs_US',
             'SCHEMA_FDR_US')
var_types<-list('mid_REVEL50to75_Miss', 'benign_REVEL50_Miss')
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
# Write out results 
write.csv(results, 'results_g_regressions_AC5orless_lessdelmiss_diagnosesrm_incSE_Eurancs.csv')
system('dx upload results_g_regressions_AC5orless_lessdelmiss_diagnosesrm_incSE_Eurancs.csv --destination ./Variant_counts/Gene_sets/')
```

### AC6to10
```{r}
# Gene set counts
AC='6to10'
system('dx download ./Variant_counts/Gene_sets/*_AC6to10_*.csv --overwrite')
file_list <- list.files(path = "./", pattern = "*_AC6to10_counts.csv", full.names = TRUE)
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames
all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

# SET UP
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC6to10_exome_wide_total + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' # SSA removed here as no one defined as EUR anc has a non-zero SSA probability. 
covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE)


# REGRESS
# Non-synonymous vars 
sets <- list('pLI_0.9','non_constrained_pLI_lessthan0.9',
             'brain_expressed','non_brain_expressed',
             'SZ_GWAS_all', 'SZ_GWAS_closest', 'SZ_GWAS_credible_causal_US',
             'NDD_CNVs_US', 'SZ_CNVs_US',
             'SCHEMA_FDR_US')
var_types<-list('PTV', 'REVEL75_Miss')
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


# Synonymous vars  
var_types<-list('synonymous')
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

# Write out results 
write.csv(results, 'results_g_regressions_AC6to10_diagnosesrm_incSE_Eurancs.csv')
system('dx upload results_g_regressions_AC6to10_diagnosesrm_incSE_Eurancs.csv --destination ./Variant_counts/Gene_sets/')
```






### AC11to20
```{r}
# Gene set counts
AC='11to20'
system('dx download ./Variant_counts/Gene_sets/*_AC11to20_*.csv --overwrite')
file_list <- list.files(path = "./", pattern = "*_AC11to20_counts.csv", full.names = TRUE)
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames
all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

# SET UP
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC11to20_exome_wide_total + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' # SSA removed here as no one defined as EUR anc has a non-zero SSA probability. 
covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE)


# REGRESS
# Non-synonymous vars 
sets <- list('pLI_0.9','non_constrained_pLI_lessthan0.9',
             'brain_expressed','non_brain_expressed',
             'SZ_GWAS_all', 'SZ_GWAS_closest', 'SZ_GWAS_credible_causal_US',
             'NDD_CNVs_US', 'SZ_CNVs_US',
             'SCHEMA_FDR_US')
var_types<-list('PTV', 'REVEL75_Miss')
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


# Synonymous vars  
var_types<-list('synonymous')
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

# Write out results 
write.csv(results, 'results_g_regressions_AC11to20_diagnosesrm_incSE_Eurancs.csv')
system('dx upload results_g_regressions_AC11to20_diagnosesrm_incSE_Eurancs.csv --destination ./Variant_counts/Gene_sets/')
```






### AC21to50
```{r}
# Gene set counts
AC='21to50'
system('dx download ./Variant_counts/Gene_sets/*_AC21to50_*.csv --overwrite')
file_list <- list.files(path = "./", pattern = "*_AC21to50_counts.csv", full.names = TRUE)
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames
all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8) 

# SET UP
covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC21to50_exome_wide_total + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' # SSA removed here as no one defined as EUR anc has a non-zero SSA probability. 
covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + AAC + EAS + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE)


# REGRESS
# Non-synonymous vars 
sets <- list('pLI_0.9','non_constrained_pLI_lessthan0.9',
             'brain_expressed','non_brain_expressed',
             'SZ_GWAS_all', 'SZ_GWAS_closest', 'SZ_GWAS_credible_causal_US',
             'NDD_CNVs_US', 'SZ_CNVs_US',
             'SCHEMA_FDR_US')
var_types<-list('PTV', 'REVEL75_Miss')
dependent_var<-list('g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


# Synonymous vars  
var_types<-list('synonymous')
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs I need
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EUR_only'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

# Write out results 
write.csv(results, 'results_g_regressions_AC21to50_diagnosesrm_incSE_Eurancs.csv')
system('dx upload results_g_regressions_AC21to50_diagnosesrm_incSE_Eurancs.csv --destination ./Variant_counts/Gene_sets/')
```




