---
title: "20_plot_results"
output: html_document
date: "2024-10-18"
---

# Plot results
All results of regressions were downloaded from the RAP and results were plotted locally. 

# Load libraries
```{r}
library(ggplot2)
library(dplyr)
y_title <- expression(paste("Beta of association with ", italic("g")))
```
# Read in results 
```{r}
results=read.csv('/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/results_g_regressions_AC5orless_diagnosesrm_incSE_crossancs_Feb25.csv')

# Drop first col
results <- results %>% select(-1)

results <- results %>%
  mutate(ann = case_when(
    grepl("PTV", indep_var) ~ "PTV",
    grepl("REVEL75_", indep_var) ~ "Deleterious missense",
    grepl("synon", indep_var) ~ "Synonymous",
    TRUE ~ NA_character_
  )) %>%
  mutate(ann = factor(ann, levels = c("PTV", 
                                      "Deleterious missense",
                                      "Synonymous")))

results %>% count(ancestry)
all_results=results

main_sets_eur_results=subset(all_results, all_results$ancestry=='EUR_only')
main_sets_eur_results=subset(main_sets_eur_results, main_sets_eur_results$dep_var=='g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

lofi_results=subset(all_results, all_results$gene_set=='pLI_0.9')
lofi_results=subset(lofi_results, lofi_results$dep_var=='g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')

Anc_specific_lofi_results=subset(lofi_results, lofi_results$ancestry=='ADM_ancs' |
                                   lofi_results$ancestry=='EAS_ancs' |
                                   lofi_results$ancestry=='SAS_ancs' |
                                   lofi_results$ancestry=='SSA_ancs' |
                                   lofi_results$ancestry=='EUR_only' |
                                   lofi_results$ancestry=='AAC_ancs' |
                                   lofi_results$ancestry=='LAT_ancs')

sex_strat_lofi_results=subset(lofi_results, lofi_results$ancestry=='EUR_only_females' |
                                lofi_results$ancestry=='EUR_only' |
                                lofi_results$ancestry=='EUR_only_males')

cognitive_covars=subset(all_results, 
                        all_results$ancestry=='EUR_only_covar_FI_AC1_n_correct_answers_Zscore' | 
                          all_results$ancestry=='EUR_only_covar_Numeric_memory_online_maximum_digits_remembered_Zscore' | 
                          all_results$ancestry=='EUR_only_covar_Pairs_matching_AC1_n_incorrect_matches_Z_score_outliersrm' | 
                          all_results$ancestry=='EUR_only_covar_Reaction_time_AC1_mean_time_taken_Zscore_outliersrm' | 
                          all_results$ancestry=='EUR_only_covar_SDS_online1_n_correct_cons_attempts_Zscore_outliersrm' | 
                          all_results$ancestry=='EUR_only_covar_TMTA_online1_time_to_complete_numeric_path_Zscore_outliersrm' | 
                          all_results$ancestry=='EUR_only_covar_TMTB_online1_time_to_complete_numeric_path_Zscore_outliersrm' | 
                          all_results$ancestry=='EUR_only_covar_g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm' |
                          (all_results$ancestry=='EUR_only' & all_results$dep_var!='g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm'))


```


# Supplementary Figure 1 - main findings summary
```{r}
gg=subset(main_sets_eur_results, main_sets_eur_results$gene_set=='pLI_0.9'| 
            main_sets_eur_results$gene_set=='SZ_GWAS_credible_causal_US' |
            main_sets_eur_results$gene_set=='SCHEMA_FDR_US')
gg$Set[gg$gene_set=='pLI_0.9']<-'Constrained genes' 
gg$Set[gg$gene_set=='SZ_GWAS_credible_causal_US']<-'Credible causal \n schizophrenia GWAS genes' 
gg$Set[gg$gene_set=='SCHEMA_FDR_US']<-'Genes enriched for rare coding \n  variants in schizophrenia' 

gg=subset(gg, gg$ann != 'Synonymous')

gg<- gg %>% mutate(Set = factor(Set, levels = c('Credible causal \n schizophrenia GWAS genes',
                                                'Genes enriched for rare coding \n  variants in schizophrenia', 
                                                'Constrained genes' )))

tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/main_findings_SF1.tiff", units="in", width=18, height=5, res=400)
ggplot(gg, aes(y=Set, x=beta, colour = ann, group=ann, shape=ann))  +
  geom_pointrange(aes(xmin=lower_CI, xmax=upper_CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_vline(xintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=30))+
  labs(x=y_title, y="Gene set", colour = 'Rare variant class', shape = 'Rare variant class')+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20))
dev.off()


```

# Constraint
## Figure 1 - eur ancs 
```{r}
# Lofi and lof tol in one plot
gg=subset(main_sets_eur_results, main_sets_eur_results$gene_set=='non_constrained_pLI_lessthan0.9'| main_sets_eur_results$gene_set=='pLI_0.9')
gg$Set[gg$gene_set=='non_constrained_pLI_lessthan0.9']<-'LoF-tolerant genes' 
gg$Set[gg$gene_set=='pLI_0.9']<-'LoF-intolerant genes' 



tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/LoF-tolerantandintolerant_AC5orless_assocs_EU_ancs.tiff", units="in", width=12, height=10, res=400)
ggplot(gg, aes(y=beta, x=Set, colour = ann, group=ann, shape=ann))  +
  geom_pointrange(aes(ymin=lower_CI, ymax=upper_CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Gene set", colour = 'Rare variant class', shape = 'Rare variant class')+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20))
dev.off()

# Z test comparing pathogenic missense in LoFi to pathogenic missense in LoF-tol
gg=subset(gg, gg$ann=="Deleterious missense")
gg$z_val<-(((-0.03982036) - gg$beta)/sqrt((0.006854876)^2 + (gg$SE)^2))
gg$z_test_p<-1-pnorm(abs(gg$z_val),0,1) # Chec the one for revel.75 ==0.5
```

## Sex strat - supp figure 4
```{r}
gg=subset(sex_strat_lofi_results, sex_strat_lofi_results$gene_set=='pLI_0.9')
gg$sex[gg$ancestry=='EUR_only']<-'Both sexes'
gg$sex[gg$ancestry=='EUR_only_males']<-'Males'
gg$sex[gg$ancestry=='EUR_only_females']<-'Females'

tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/constraint_AC5orless_assocs_EU_ancs_sexstrat.tiff", units="in", width=12, height=10, res=400)
ggplot(gg, aes(y=beta, x=ann, colour = ann, group=sex, shape=sex))  +
  geom_pointrange(aes(ymin=lower_CI, ymax=upper_CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Variant class")+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20))+
  guides(colour=guide_legend(title=NULL))  
dev.off()

```
## Brain expression (do we want a section on this?)
```{r}
# Lofi and lof tol in one plot
gg=subset(main_sets_eur_results, main_sets_eur_results$gene_set=='brain_expressed'| main_sets_eur_results$gene_set=='non_brain_expressed')
gg$Set[gg$gene_set=='brain_expressed']<-'Brain expressed genes' 
gg$Set[gg$gene_set=='non_brain_expressed']<-'Non-brain expressed genes' 



tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/LoF-tolerantandintolerant_AC5orless_assocs_EU_ancs.tiff", units="in", width=12, height=10, res=400)
ggplot(gg, aes(y=beta, x=Set, colour = ann, group=ann, shape=ann))  +
  geom_pointrange(aes(ymin=lower_CI, ymax=upper_CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Gene set", colour = 'Rare variant class', shape = 'Rare variant class')+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20))
dev.off()

```

## Brain expression and lof tolerance
```{r}
gg=subset(main_sets_eur_results, main_sets_eur_results$gene_set=='brain_expressed_LoFi'| 
            main_sets_eur_results$gene_set== 'non_brain_expressed_LoFi')

# brain expressed lofi vs non brain expressed lofi
# miss
z=subset(gg, gg$ann=="Deleterious missense")
z$z_val<-(((-0.03959504) - z$beta)/sqrt((0.007813874)^2 + (z$SE)^2))
z$z_test_p<-1-pnorm(abs(z$z_val),0,1) # Chec the one for revel.75 ==0.5

# PTVs
z=subset(gg, gg$ann=="PTV")
z$z_val<-(((-0.07023037) - z$beta)/sqrt((0.008105364)^2 + (z$SE)^2))
z$z_test_p<-1-pnorm(abs(z$z_val),0,1) # Chec the one for revel.75 ==0.5


gg=subset(main_sets_eur_results, main_sets_eur_results$gene_set=='brain_expressed_LoFtolerant'| 
            main_sets_eur_results$gene_set== 'non_brain_expressed_LoFtolerant')

```

## Covarying for individual cognitive tests
### When g is outcome measure
```{r}
gg=subset(cognitive_covars, cognitive_covars$dep_var=='g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')
gg=subset(gg, gg$ann!='Synonymous')

```

### Other outcome measures
```{r}
gg=read.csv('/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/results_diff_outcomes_covarg_regressions_AC5orless.csv')
subset(cognitive_covars, cognitive_covars$dep_var!='g_MAIN_RT_TMTB_NM_PM_zscore_outliersrm')
gg=subset(gg, gg$indep_var!='synonymous_AC5orless_pLI_0.9_total')

```



## SF5 - LoFi across ancestries 
```{r}
# Lofi and lof tol in one plot
gg=Anc_specific_lofi_results
gg$Genetic_ancestry[gg$ancestry=="AAC_ancs"]<-'1KGP-AAC-like \n (305 people \n with g)'
gg$Genetic_ancestry[gg$ancestry=="ADM_ancs"]<-'Admixed \n (92 people \n with g)'
gg$Genetic_ancestry[gg$ancestry=="SAS_ancs"]<-'1KGP-SAS-like \n (665 people \n with g)'
gg$Genetic_ancestry[gg$ancestry=="SSA_ancs"]<-'1KGP-SSA-like \n (147 people \n with g)'
gg$Genetic_ancestry[gg$ancestry=="EAS_ancs"]<-'1KGP-EAS-like \n (293 people \n with g)'
gg$Genetic_ancestry[gg$ancestry=="LAT_ancs"]<-'1KGP-LAT-like \n (93 people \n with g)'
gg$Genetic_ancestry[gg$ancestry=="EUR_only"]<-'1KGP-EUR-like \n (75,188 people \n with g)'

gg<- gg %>% mutate(Genetic_ancestry = factor(Genetic_ancestry, levels = c('1KGP-EUR-like \n (75,188 people \n with g)',
                                                                          '1KGP-SAS-like \n (665 people \n with g)',
                                                                          '1KGP-AAC-like \n (305 people \n with g)',
                                                                          '1KGP-EAS-like \n (293 people \n with g)',
                                                                          '1KGP-SSA-like \n (147 people \n with g)',
                                                                          '1KGP-LAT-like \n (93 people \n with g)',
                                                                          'Admixed \n (92 people \n with g)')))




tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/LoFi_AC5orless_across_ancs.tiff", units="in", width=12, height=10, res=400)
ggplot(gg, aes(y=beta, x=Genetic_ancestry, colour = ann, group=ann, shape=ann))  +
  geom_pointrange(aes(ymin=lower_CI, ymax=upper_CI), position = position_dodge(width = 0.55), size=1.8)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Genetic ancestry", colour = 'Rare variant class', shape = 'Rare variant class')+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20),
        axis.text.x = element_text(size = 15))
dev.off()

# Z tests
gg=subset(gg, gg$ann=="PTV")
gg$z_val<-(((-0.0622877047) - gg$beta)/sqrt((0.007383138)^2 + (gg$SE)^2))
gg$z_test_p<-1-pnorm(abs(gg$z_val),0,1) # Chec the one for revel.75 ==0.5

gg=Anc_specific_lofi_results
gg=subset(gg, gg$ann=="Deleterious missense")
gg$z_val<-(((-0.039820357) - gg$beta)/sqrt((0.006854876)^2 + (gg$SE)^2))
gg$z_test_p<-1-pnorm(abs(gg$z_val),0,1) # Chec the one for revel.75 ==0.5
```

# Figure 2 - SCHEMA FDR genes 
```{r}
gg=subset(results, results$gene_set=='SCHEMA_FDR_US')
gg$Set<-'SCHEMA FDR < 5% genes'

tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/SCHEMA_FDR_AC55orless_assocs_eur_ancs.tiff", units="in", width=12, height=10, res=400)
ggplot(gg, aes(y=beta, x=Set, colour = ann, group=ann, shape=ann))  +
  geom_pointrange(aes(ymin=lower_CI, ymax=upper_CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Gene set", colour = 'Rare variant class', shape = 'Rare variant class')+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20))
dev.off()

# And z test 
gg=subset(results, results$gene_set=='SCHEMA_FDR_US'|  results$gene_set=='LoFi_minus_SCHEMA_FDR_US')
z=subset(gg, gg$ann=="PTV")
z$z_val<-(((-0.24488849) - z$beta)/sqrt((0.068588162)^2 + (z$SE)^2))
z$z_test_p<-1-pnorm(abs(z$z_val),0,1) # Chec the one for revel.75 ==0.5
```
# Figure 4 -  GWAS genes
```{r}
# three sets in one plot
gg=subset(results, results$gene_set=='SZ_GWAS_closest'| results$gene_set=='SZ_GWAS_all' | results$gene_set=='SZ_GWAS_credible_causal_US')

gg$Set[gg$gene_set=='SZ_GWAS_closest']<-'Genes nearest to the \n locus index SNPs in \n the SZ GWAS' 
gg$Set[gg$gene_set=='SZ_GWAS_all']<-'All genes within \n SZ GWAS loci'
gg$Set[gg$gene_set=='SZ_GWAS_credible_causal_US']<-'Genes prioritised as \n credible-causal in the \n SZ GWAS'

gg<- gg %>% mutate(Set = factor(Set, levels = c('All genes within \n SZ GWAS loci', 
                                                'Genes nearest to the \n locus index SNPs in \n the SZ GWAS' , 
                                                'Genes prioritised as \n credible-causal in the \n SZ GWAS')))


tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/GWAS_sets_AC5orless_assocs_EU_ancs.tiff", units="in", width=12, height=10, res=400)
ggplot(gg, aes(y=beta, x=Set, colour = ann, group=ann, shape=ann))  +
  geom_pointrange(aes(ymin=lower_CI, ymax=upper_CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Gene set", colour = 'Rare variant class', shape = 'Rare variant class')+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20))
dev.off()

# Additional sets
gg=subset(results, results$gene_set=='non_prioritised_GWAS'| results$gene_set=='Finemap_prioritised_GWAS' | results$gene_set=='SMR_prioritised_GWAS' | results$gene_set=='SZ_GWAS_closest'| results$gene_set=='SZ_GWAS_all' | results$gene_set=='SZ_GWAS_credible_causal_US')
gg<- gg %>% mutate(gene_set = factor(gene_set, levels = c('non_prioritised_GWAS', 
                                                'Finemap_prioritised_GWAS' , 
                                                'SMR_prioritised_GWAS')))

# Comparisons
gg=subset(results, results$gene_set=='non_prioritised_GWAS'| results$gene_set=='Finemap_prioritised_GWAS' | results$gene_set=='SMR_prioritised_GWAS' | results$gene_set=='SZ_GWAS_closest'| results$gene_set=='SZ_GWAS_all' | results$gene_set=='SZ_GWAS_credible_causal_US')
# And z tests
gg=subset(results, results$gene_set=='SZ_GWAS_credible_causal_US'|  results$gene_set=='non_prioritised_GWAS')
z=subset(gg, gg$ann=="Deleterious missense")
z$z_val<-(((-0.118096306) - z$beta)/sqrt((0.03392730)^2 + (z$SE)^2))
z$z_test_p<-1-pnorm(abs(z$z_val),0,1) # Chec the one for revel.75 ==0.5

gg=subset(results, results$gene_set=='SZ_GWAS_credible_causal_US'|  results$gene_set=='LoFi_minus_credible_causal_GWAS')
z=subset(gg, gg$ann=="Deleterious missense")
z$z_val<-(((-0.118096306) - z$beta)/sqrt((0.03392730)^2 + (z$SE)^2))
z$z_test_p<-1-pnorm(abs(z$z_val),0,1) # Chec the one for revel.75 ==0.5

gg=subset(results, results$gene_set=='SZ_GWAS_credible_causal_US'|  results$gene_set=='Brain_expressed_minus_credible_causal_GWAS')
z=subset(gg, gg$ann=="Deleterious missense")
z$z_val<-(((-0.118096306) - z$beta)/sqrt((0.03392730)^2 + (z$SE)^2))
z$z_test_p<-1-pnorm(abs(z$z_val),0,1) # Chec the one for revel.75 ==0.5

gg=subset(results, results$gene_set=='SMR_prioritised_GWAS'|  results$gene_set=='Brain_expressed_minus_SMR_prioritised_GWAS')
z=subset(gg, gg$ann=="Deleterious missense")
z$z_val<-(((-0.11816858) - z$beta)/sqrt((0.05911149)^2 + (z$SE)^2))
z$z_test_p<-1-pnorm(abs(z$z_val),0,1) # Chec the one for revel.75 ==0.5
```

# Supplementary FIgure 6
```{r}
gg=subset(results, results$gene_set=='SZ_GWAS_credible_causal_US'| 
            results$gene_set=='Finemap_prioritised_GWAS' | 
            results$gene_set=='SMR_prioritised_GWAS')
gg$Set[gg$gene_set=='Finemap_prioritised_GWAS']<-'Finemap prioritised \n genes' 
gg$Set[gg$gene_set=='SMR_prioritised_GWAS']<-'SMR prioritised \n genes'
gg$Set[gg$gene_set=='SZ_GWAS_credible_causal_US']<-'Genes prioritised as \n credible-causal in the \n SZ GWAS'

gg<- gg %>% mutate(Set = factor(Set, levels = c( 'Genes prioritised as \n credible-causal in the \n SZ GWAS', 
                                                 'Finemap prioritised \n genes',
                                                 'SMR prioritised \n genes')))
                                            

tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/GWAS_diffprior_sets_AC5orless_assocs_EU_ancs.tiff", units="in", width=12, height=10, res=400)
ggplot(gg, aes(y=beta, x=Set, colour = ann, group=ann, shape=ann))  +
  geom_pointrange(aes(ymin=lower_CI, ymax=upper_CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Gene set", colour = 'Rare variant class', shape = 'Rare variant class')+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20))
dev.off()

```


# Figure 5 - CNVs
```{r}
# three sets in one plot
gg=subset(results, results$gene_set=='NDD_CNVs_US' | results$gene_set=='LoFi_in_NDD_CNVs_US' | results$gene_set=='LoFtolerant_in_NDD_CNVs_US')
          
gg$Set[gg$gene_set=='NDD_CNVs_US']<-'All genes within \n schizophrenia-enriched \n CNV loci' 
gg$Set[gg$gene_set=='LoFi_in_NDD_CNVs_US']<-'LoFi genes within \n schizophrenia-enriched \n CNV loci'
gg$Set[gg$gene_set=='LoFtolerant_in_NDD_CNVs_US']<-'LoF-tolerant genes \n within schizophrenia- \n enriched CNV loci'

gg<- gg %>% mutate(Set = factor(Set, levels = c('All genes within \n schizophrenia-enriched \n CNV loci',
                                                'LoFi genes within \n schizophrenia-enriched \n CNV loci',
                                                'LoF-tolerant genes \n within schizophrenia- \n enriched CNV loci')))


tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/szenrichedCNV_sets_AC5orless_assocs_EU_ancs.tiff", units="in", width=12, height=10, res=400)
ggplot(gg, aes(y=beta, x=Set, colour = ann, group=ann, shape=ann))  +
  geom_pointrange(aes(ymin=lower_CI, ymax=upper_CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Gene set", colour = 'Rare variant class', shape = 'Rare variant class')+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20))
dev.off()

# And z tests
gg=subset(results, results$gene_set=='LoFi_in_NDD_CNVs_US'|  results$gene_set=='LoFi_minus_NDD_CNVs_US')
z=subset(gg, gg$ann=="PTV")
z$z_val<-(((-0.08648960) - z$beta)/sqrt((0.033538832)^2 + (z$SE)^2))
z$z_test_p<-1-pnorm(abs(z$z_val),0,1) # Chec the one for revel.75 ==0.5


```

## SZ CNVs
```{r}
# three sets in one plot
gg=subset(results, results$gene_set=='SZ_CNVs_US' | results$gene_set=='LoFi_in_SZ_CNVs_US' | results$gene_set=='LoFtolerant_in_SZ_CNVs_US')
          
gg$Set[gg$gene_set=='SZ_CNVs_US']<-'All genes within \n schizophrenia-associated \n CNV loci' 
gg$Set[gg$gene_set=='LoFi_in_SZ_CNVs_US']<-'LoFi genes within \n schizophrenia-associated \n CNV loci'
gg$Set[gg$gene_set=='LoFtolerant_in_SZ_CNVs_US']<-'LoF-tolerant genes \n within schizophrenia- \n associated CNV loci'

gg<- gg %>% mutate(Set = factor(Set, levels = c('All genes within \n schizophrenia-associated \n CNV loci',
                                                'LoFi genes within \n schizophrenia-associated \n CNV loci',
                                                'LoF-tolerant genes \n within schizophrenia- \n associated CNV loci')))


```

## Single gene burden tests
```{r}
ptv=read.csv('/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/PTV_EUR_anc_AC5orless_single_gene_burden.csv')
ptv$variant<-'Rare PTVs'
miss=read.csv('/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/REVEL75_Miss_EUR_anc_AC5orless_single_gene_burden.csv')
miss$variant<-'Rare pathogenic missense variants'
both=read.csv('/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/PTVs_and_Miss_EUR_anc_AC5orless_single_gene_burden.csv')
both$variant<-'Rare PTVs + pathogenic missense'
gene_annotations=read.csv('/Users/c1928239/Downloads/gene_sets_for_RAP_Feb25.csv')

# Some of these are dups of gene ids 
duplicates <- gene_annotations %>%
  group_by(gene_id) %>%
  filter(n() > 1)
# remove them 
gene_annotations_unique <- gene_annotations %>%
  group_by(gene_id) %>%
  slice(1) %>%  # Keeps only the first row for each gene_id
  ungroup()

gene_annotations_unique %>% count(pLI_constrained)
gene_annotations_unique %>% count(Brain_expressed)
gene_annotations_unique %>% count()

```

```{r}
single_gene=rbind(ptv, miss, both)
all_single_gene<-merge(single_gene, gene_annotations_unique, by.x='gene_set', by.y='gene_id', all.x=T)
all_single_gene<-all_single_gene[,c(16,1,14,21,35,46,53,54,63,6,7,8,9,10,11)]


all_single_gene$SZ_GWAS_credible_causal[is.na(all_single_gene$SZ_GWAS_credible_causal)]<-F
all_single_gene$NDD_CNV_gene[is.na(all_single_gene$NDD_CNV_gene)]<-F
all_single_gene$SZ_CNV_gene[is.na(all_single_gene$SZ_CNV_gene)]<-F
all_single_gene$SCHEMA_FDR5_gene[is.na(all_single_gene$SCHEMA_FDR5_gene)]<-F

write.csv(all_single_gene, '/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/MAC5_single_gene_results.csv', row.names = F)

all_single_gene_5andmore=all_single_gene
all_single_gene_5andmore$beta[all_single_gene_5andmore$n_vars<5]<-NA
all_single_gene_5andmore$SE[all_single_gene_5andmore$n_vars<5]<-NA
all_single_gene_5andmore$lower_CI[all_single_gene_5andmore$n_vars<5]<-NA
all_single_gene_5andmore$upper_CI[all_single_gene_5andmore$n_vars<5]<-NA
all_single_gene_5andmore$p_value[all_single_gene_5andmore$n_vars<5]<-NA
all_single_gene_5andmore$n_vars[all_single_gene_5andmore$n_vars<5]<-'<5'
write.csv(all_single_gene_5andmore, '/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/MAC5_single_gene_results_fewerthan5rm.csv', row.names = F)
```


# Combined analysis
```{r}
univar<-read.csv('/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/univariate_comparison_betas.csv')

univar %>% count(Variant.type)
univar$Variant_type[univar$Variant.type=="Rare PTVs in LoFi genes"]<- "Rare PTVs in \n LoFi genes"
univar$Variant_type[univar$Variant.type=="Rare pathogenic missense variants in LoFi genes "]<- "Rare deleterious \n missense variants \n in LoFi genes"
univar$Variant_type[univar$Variant.type=="Schizophrenia PRS"]<- "Schizophrenia \n PRS"
univar$Variant_type[univar$Variant.type=="Schizophrenia enriched  CNVs"]<-"Schizophrenia \n enriched CNVs"      
univar$Variant_type[univar$Variant.type=="Rare PTVs in SCHEMA FDR<5% genes"]<-"Rare PTVs \n in SCHEMA FDR \n <5% genes"
univar$Variant_type[univar$Variant.type=="Rare deleterious missense variants in credoble causa schizophrenia GWAS genes"]<-"Rare deleterious \n missense variants in \n credible causal \n schizophrenia GWAS genes"
univar$Variant_type[univar$Variant.type=="Rare PTVs in LoFi genes within schizophrenia enriched CNVs"]<-"Rare PTVs in \n LoFi genes within \n schizophrenia enriched \n CNVs"
univar$Variant_type[univar$Beta== -0.086]<-"Rare PTVs in \n LoFi genes within \n schizophrenia enriched \n CNVs"

univar <- univar %>% mutate(Variant_type = factor(Variant_type, levels = c("Rare deleterious \n missense variants \n in LoFi genes",
                                                "Rare PTVs in \n LoFi genes",
                                                "Rare PTVs in \n LoFi genes within \n schizophrenia enriched \n CNVs",
                                                "Rare deleterious \n missense variants in \n credible causal \n schizophrenia GWAS genes",
                                                "Rare PTVs \n in SCHEMA FDR \n <5% genes",
                                                "Schizophrenia \n enriched CNVs",
                                                "Schizophrenia \n PRS")))
                                                


tiff("/Users/c1928239/OneDrive - Cardiff University/Cardiff/UKBB RAP work/diff_var_classes_betas_EU_ancs.tiff", units="in", width=12, height=10, res=400)
ggplot(univar, aes(y=Beta, x=Variant_type))  +
  geom_pointrange(aes(ymin=Lower.CI, ymax=Upper.CI), position = position_dodge(width = 0.55), size=2.5)+
  geom_hline(yintercept=0, linetype="dotted")+
  theme_light()+
  theme(text=element_text(size=25))+
  labs(y=y_title, x="\n Variant type")+
  theme(legend.position="bottom")+
  theme(legend.text = element_text(size=20),
        axis.text.x = element_text(size=10))
dev.off()

```



